---
title: "manuscript_Ahlqvist_analysis"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
Although diabetes is broadly classified as type 1 and type 2 diabetes, type 2 diabetes is known to be heterogeneous. Finer classifications of diabetes could aid in the personalization of diabetes treatment. [Ahlqvist et al.](https://doi.org/10.1016/s2213-8587(18)30051-2) recently identified 5 subtypes of diabetes with different risks of diabetic complications using glutamate decarboxylase antibody status, age at diagnoses, body mass index, HbA1c, and homeostatic model assessment 2 estimates of beta-cell function and insulin resistance. 

**Objective:** Conduct statistical analysis and generate data to construct figures in `manuscript_Ahlqvist_figures.ipynb`.

# Environment set up

```{r}
library(RSQLite)
library(tidyverse)
library(car)
library(lme4)
library(lmerTest)
library(openxlsx)
```

```{r}
import_dir <- "/data/user/brianlu/projects/precision_diabetes/COVID19_clusters/exports"
post_covid_path <- file.path(import_dir, "remap_clusters_rf_2024-07-19/remapped_outpatient.csv")

post_covid <- read_csv(post_covid_path)
```

# Feature engineering

## Clean up post-COVID19 data to match pre-COVID19 data
Add units.

```{r}
post_covid <- post_covid %>%
  mutate(homa_cpeptide_b_units = "%",
         homa_cpeptide_s_units = "%",
         homa_cpeptide_ir_units = "Index",
         homa_insulin_b_units = "%",
         homa_insulin_s_units = "%",
         homa_insulin_ir_units = "Index",
         diabetes_age_units = "Years",
         bmi_units = "kg/m^2")
```

Recode gender from just first letter abbreviations to full name.

```{r}
post_covid <- post_covid %>%
  rename(gender = GENDER) %>%
  mutate(gender = case_match(gender,
                             "M" ~ "Male",
                             "F" ~ "Female"))
```

Recode first positive urine ketone from string to postive or negative.

```{r}
post_covid$u_ketone_first_positive <- ifelse(is.na(post_covid$u_ketone_first_positive), 0, 1)
```

Rename DECEASED to deceased.

```{r}
post_covid <- post_covid %>% rename(deceased = DECEASED)
```

There were too few patients who are not White or Black or African American for meaningful analysis. Combine races other than White or Black/African American into Other.

```{r}
post_covid <- post_covid %>% 
  rename(race_original = RACE) %>% 
  mutate(race = ifelse(race_original == "White" | race_original == "Black or African American", race_original, "Other"))
```

Calculate triglyceride/hdl and ldl/hdl ratios.

```{r}
post_covid <- post_covid %>% mutate(triglyceride_hdl = ifelse(.$triglyceride_date == .$hdl_date, .$triglyceride/.$hdl, NA),
                                      triglyceride_hdl_units = "Ratio",
                                      ldl_hdl = ifelse(.$ldl_date == .$hdl_date, .$ldl/.$hdl, NA),
                                      ldl_hdl_units = "Ratio")

```

Classify BMI.

```{r}
post_covid$bmi_class <- ifelse(post_covid$race_original == "Asian" & post_covid$bmi < 18.5, "underweight",
                    ifelse(post_covid$race_original == "Asian" & post_covid$bmi >= 18.5 & post_covid$bmi < 23, "normal",
                           ifelse(post_covid$race_original == "Asian" & post_covid$bmi >= 23 & post_covid$bmi < 27.5, "overweight",
                                  ifelse(post_covid$race_original == "Asian" & post_covid$bmi >= 27.5, "obese",
                                         
                                         ifelse(post_covid$race_original != "Asian" & post_covid$bmi <= 18.5, "underweight",
                                                ifelse(post_covid$race_original != "Asian" & post_covid$bmi >= 18.5 & post_covid$bmi < 25, "normal",
                                                       ifelse(post_covid$race_original != "Asian" & post_covid$bmi >= 25 & post_covid$bmi < 30, "overweight",
                                                              ifelse(post_covid$race_original != "Asian" & post_covid$bmi >= 30, "obese", NA))))))))

```

Label each Ahlqvist cluster for leave-one-out comparisons.

```{r}
order <- c("SAID", "SIDD", "SIRD", "MOD", "MARD")

for (c in order){
    post_covid[[c]] = ifelse(post_covid$mapped_cluster == c, 1, 0)
}
```

Calculate deceased age.

```{r}
post_covid <- post_covid %>%
  mutate(deceased_age = as.numeric(deceased_date - birth_date, "days")/365,
         deceased_age_units = "Years")
```

Rename cluster.

```{r}
post_covid <- post_covid %>% rename(cluster = mapped_cluster) 

```

Label post-COVID19 and insert censor date.

```{r}
# find censor date
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2024-06-17.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

covid_censored_date <- dbGetQuery(conn, "SELECT DISTINCT DIAG_DT_TM FROM diagnoses") %>%
  as_tibble() %>%
  mutate(date = date(mdy_hms(DIAG_DT_TM))) %>%
  pull(date) %>%
  max()

dbDisconnect(conn)

# insert label and date
post_covid <- post_covid %>%
  mutate(pandemic = "post",
         covid19_censored_date = case_when(!is.na(covid19_date) ~ covid19_date,
                                           .default = ymd(covid_censored_date)))



```

Find deceased censoring date for pre-COVID data. Last diagnosis date is 2020-01-01, but deceased date extends past 2020-01-01, probably to when data was pulled. The last deceased date for post-COVID data is prior to the last diagnosis date.

```{r}
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2023-06-26.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

deceased_censor_date <- dbGetQuery(conn, "SELECT DISTINCT DECEASED_DATE FROM patient_demographics") %>%
    as_tibble() %>%
    mutate(date = date(mdy_hms(DECEASED_DATE))) %>%
    pull(date) %>%
    max(na.rm = T)

dbDisconnect(conn)
```

Retrieve insulin prescription data.

```{r}
# connect to SQL database
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2024-06-17.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

# get matching MRNS for DE_ID in SIDD White and AA patients
mrn <- dbGetQuery(conn, "SELECT MRN, DE_ID FROM mrns") %>% 
  as_tibble() %>% 
  filter(DE_ID %in% post_covid$DE_ID)

# search medication records by MRNS
years <- c(seq(2020, 2024))

# add single quotes to beginning and end of string
mrn_list <- str_replace_all(as.character(mrn$MRN), "^(.*)$", "'\\1'")

# declare list to aggregate prescription data
prescription_list <- list()

prescription_names <- c("insulin", 
                        #"glucagon", 
                        "metformin", 
                        #"sulfonylurea", 
                        #"SGLT2i"
                        "GLP1RA"
                        )
prescription_search <- c("'%insulin%'", 
                         #"'%glucagon%'",
                         "'%metformin%'", 
                         #"'%glimepiride%' or ORDERABLE LIKE '%glipizide%' or ORDERABLE LIKE '%glyburide%'",
                         #"'%bexagliflozin%' or ORDERABLE LIKE '%canagliflozin%' or ORDERABLE LIKE '%dapagliflozin%' or ORDERABLE LIKE '%empagliflozin%'"
                         "'%dulaglutide%' or ORDERABLE LIKE '%albiglutide%' or ORDERABLE LIKE  '%liraglutide%' or ORDERABLE LIKE '%semaglutide%' or ORDERABLE LIKE '%exenatide%' or ORDERABLE LIKE '%lixisenatide%' or ORDERABLE LIKE '%tirzepatide%'"
                         )
names(prescription_search) <- prescription_names

for (p in prescription_names){
  # aggregate prescription data
  years_list <- list()
  print(paste("Querying", p))
  
  for (y in years){
    print(paste("Retrieving", y, "records"))
    
    table_name <- paste("all_meds", y, sep = "_")
    years_list[[paste("year", y, sep = "_")]] <- dbGetQuery(conn, paste0("SELECT MRN, ORDER_DT_TM ",
                                                              "FROM ", table_name, " ",
                                                              "WHERE (ORDERABLE LIKE ", prescription_search[p], ") and ", 
                                                              "MRN IN (", paste(mrn_list, collapse = ","), ")")) %>% 
      as_tibble() %>%
      mutate(order_date = date(mdy_hms(ORDER_DT_TM))) %>% 
      filter(order_date >= "2020-01-01") %>%
      distinct(MRN, .keep_all = T)
  
  }
  # merge list into single tibble
  prescription_list[[p]] <- bind_rows(years_list)
  
  # match to DE_ID
  prescription_list[[p]] <- mrn %>% 
    left_join(prescription_list[[p]], by = c("MRN" = "MRN")) %>% 
    select(-MRN, -ORDER_DT_TM) %>% # remove identifying MRNS
    rename("{p}_prescription_date" := order_date) %>% 
    group_by(DE_ID) %>% 
    arrange(get(paste0(p, "_prescription_date")), .by_group = T) %>% # arrange to find first prescription
    distinct(DE_ID, .keep_all = T) %>% # only keep first prescription
    mutate("{p}_prescription" := ifelse(is.na(.data[[paste0(p, "_prescription_date")]]), 0, 1)) %>% # 0 if no prescription
    mutate("{p}_prescription_date" := case_when(get(paste(p, "prescription", sep = "_")) == 0 ~ ymd(covid_censored_date), 
                                                .default = get(paste0(p, "_prescription_date")))) # insert censor date

}

# disconnect connection to SQL
dbDisconnect(conn)

# merge into 1 tibble
prescription <- prescription_list[["insulin"]]
for (p in prescription_names[prescription_names != "insulin"]){
  prescription <- prescription %>% full_join(prescription_list[[p]], by = join_by(DE_ID))
}

# match DE_ID type before joining
prescription <- prescription %>% mutate(DE_ID = as.numeric(DE_ID))

post_covid <- post_covid %>% left_join(prescription, by = join_by(DE_ID))

```

## Import pre-COVID19 data

```{r}
pre_covid_path <- file.path("/data/user/brianlu/projects/precision_diabetes/Clustering/exports", 
                            "manuscript_Ahlqvist_analysis_2024-05-03/outpatient.csv")

pre_covid <- read_csv(pre_covid_path)

pre_covid <- pre_covid %>% mutate(cluster = Ahlqvist,
                                  covid19 = 0,
                                  covid19_date = NA,
                                  covid19_censored_date = mdy("12-31-2019"),
                                  pandemic = "pre")
```

## Find hospitalization records

### Pre-COVID inpatient data

```{r}
# connect to SQL database
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2023-06-26.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

# extract DE_ID to query
precovid_id <- str_replace_all(as.character(pre_covid$DE_ID), "^(.*)$", "'\\1'")

# retrieve inpatient encounters
pre_covid_inpatient <- dbGetQuery(conn, paste0("SELECT DE_ID, ENCOUNTER_DATE
                                                FROM encounters 
                                                WHERE ENCOUNTER_TYPE_CLASS = 'Inpatient' AND
                                                DE_ID IN (", 
                                                paste(precovid_id, collapse = ","), 
                                                ")")) %>% 
  as_tibble() %>%
  # fix formatting
  mutate(DE_ID = as.numeric(DE_ID),
         inpatient_date = date(mdy_hms(ENCOUNTER_DATE))) %>%
  # join to get diabetes dates
  left_join(x = ., y = pre_covid %>% select(DE_ID, diabetes_date), by = join_by(DE_ID)) %>%
  # keep only inpatient encounters after diabetes diagnosis
  filter(inpatient_date >= diabetes_date) %>%
  # calculate time to inpatient encounter
  mutate(to_inpatient = as.numeric(inpatient_date - diabetes_date, "days")/365)

# count inpatient encounters per patient
pre_covid_inpatient_counts <- pre_covid_inpatient %>%
  group_by(DE_ID) %>%
  summarise(inpatient_visit_count = n(),
            to_inpatient = min(to_inpatient))

# join with rest of data
pre_covid <- pre_covid %>%
  left_join(pre_covid_inpatient_counts, by = join_by(DE_ID)) %>%
  # fix missing fields for patient without inpatient encounters
  mutate(inpatient_visit_count = case_when(is.na(inpatient_visit_count) ~ 0,
                                           .default = inpatient_visit_count),
         inpatient = case_when(inpatient_visit_count > 0 ~ 1,
                               .default = 0),
         to_inpatient = case_when(is.na(to_inpatient) ~ as.numeric(mdy("12-31-2019") - diabetes_date, "days")/365,
                                  .default = to_inpatient)
  )

```

### Post-COVID inpatient data

```{r}
# connect to SQL database
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2024-06-17.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

# extract DE_ID to query
postcovid_id <- str_replace_all(as.character(post_covid$DE_ID), "^(.*)$", "'\\1'")

# retrieve inpatient encounters
post_covid_inpatient <- dbGetQuery(conn, paste0("SELECT DE_ID, ENCOUNTER_DATE
                                                FROM encounters 
                                                WHERE ENCOUNTER_TYPE_CLASS = 'Inpatient' AND
                                                DE_ID IN (", 
                                                paste(postcovid_id, collapse = ","), 
                                                ")")) %>% 
  as_tibble() %>%
  # fix formatting
  mutate(DE_ID = as.numeric(DE_ID),
         inpatient_date = date(mdy_hms(ENCOUNTER_DATE))) %>%
  # join to get diabetes dates
  left_join(x = ., y = post_covid %>% select(DE_ID, diabetes_date), by = join_by(DE_ID)) %>%
  # keep only inpatient encounters after diabetes diagnosis
  filter(inpatient_date >= diabetes_date) %>%
  # calculate time to inpatient encounter
  mutate(to_inpatient = as.numeric(inpatient_date - diabetes_date, "days")/365)

# count inpatient encounters per patient
post_covid_inpatient_counts <- post_covid_inpatient %>%
  group_by(DE_ID) %>%
  summarise(inpatient_visit_count = n(),
            to_inpatient = min(to_inpatient))

# join with rest of data
post_covid <- post_covid %>%
  left_join(post_covid_inpatient_counts, by = join_by(DE_ID)) %>%
  # fix missing fields for patient without inpatient encounters
  mutate(inpatient_visit_count = case_when(is.na(inpatient_visit_count) ~ 0,
                                           .default = inpatient_visit_count),
         inpatient = case_when(inpatient_visit_count > 0 ~ 1,
                               .default = 0),
         to_inpatient = case_when(is.na(to_inpatient) ~ 
                                    as.numeric(ymd(covid_censored_date) - diabetes_date, "days")/365,
                                  .default = to_inpatient)
  )

dbDisconnect(conn)
```


## Find prescription data

```{r}
# connect to SQL database
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2024-01-10.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

# get matching MRNS for DE_ID in SIDD White and AA patients
mrn <- dbGetQuery(conn, "SELECT MRNS, DE_ID FROM mrns") %>% 
  as_tibble() %>% 
  filter(DE_ID %in% pre_covid$DE_ID)

# search medication records by MRNS
years <- c("pre2013", seq(2013, 2019))

# add single quotes to beginning and end of string
mrn_list <- str_replace_all(as.character(mrn$MRNS), "^(.*)$", "'\\1'")

# declare list to aggregate prescription data
prescription_list <- list()


for (p in prescription_names){
  # aggregate prescription data
  years_list <- list()
  print(paste("Querying", p))
  
  for (y in years){
    print(paste("Retrieving", y, "records"))
    
    table_name <- paste("all_meds", y, sep = "_")
    years_list[[paste("year", y, sep = "_")]] <- dbGetQuery(conn, paste0("SELECT MRN, ORDER_DT_TM ",
                                                              "FROM ", table_name, " ",
                                                              "WHERE (ORDERABLE LIKE ", prescription_search[p], ") and ", 
                                                              "MRN IN (", paste(mrn_list, collapse = ","), ")")) %>% 
      as_tibble() %>%
      mutate(order_date = date(mdy_hms(ORDER_DT_TM))) %>% 
      filter(order_date >= "2010-01-01" & order_date <= "2019-12-31") %>%
      distinct(MRN, .keep_all = T)
  
  }
  # merge list into single tibble
  prescription_list[[p]] <- bind_rows(years_list)
  
  # match to DE_ID
  prescription_list[[p]] <- mrn %>% 
    left_join(prescription_list[[p]], by = c("MRNS" = "MRN")) %>% 
    select(-MRNS, -ORDER_DT_TM) %>% # remove identifying MRNS
    rename("{p}_prescription_date" := order_date) %>% 
    group_by(DE_ID) %>% 
    arrange(get(paste0(p, "_prescription_date")), .by_group = T) %>% # arrange to find first prescription
    distinct(DE_ID, .keep_all = T) %>% # only keep first prescription
    mutate("{p}_prescription" := ifelse(is.na(.data[[paste0(p, "_prescription_date")]]), 0, 1)) %>% # 0 if no prescription
    mutate("{p}_prescription_date" := case_when(get(paste(p, "prescription", sep = "_")) == 0 ~ ymd("2019-12-31"), 
                                                .default = get(paste0(p, "_prescription_date")))) # insert censor date

}

# disconnect connection to SQL
dbDisconnect(conn)

# merge into 1 tibble
prescription <- prescription_list[["insulin"]]
for (p in prescription_names[prescription_names != "insulin"]){
  prescription <- prescription %>% full_join(prescription_list[[p]], by = join_by(DE_ID))
}

# match DE_ID type before joining
prescription <- prescription %>% mutate(DE_ID = as.numeric(DE_ID))

pre_covid <- pre_covid %>% 
  select(-insulin_order_date, -insulin_prescription) %>% # remove insulin from previous analysis
  left_join(prescription, by = join_by(DE_ID))

```

## Join data
Combine pre- and post-COVID19 data.

```{r}
join_feature <- colnames(post_covid)[(colnames(post_covid) %in% colnames(pre_covid))]

pre_covid_selected <- pre_covid %>% select(all_of(join_feature))
  
post_covid_selected <- post_covid %>% select(all_of(join_feature))

joined_data <- rbind(pre_covid_selected, post_covid_selected) %>% 
  mutate(covid19_date = as_date(covid19_date),
         date_range = factor(case_when(diabetes_date < "2015-01-01" ~ "y2010",
                                       diabetes_date >= "2015-01-01" & diabetes_date < "2020-01-01" ~ "y2015",
                                       diabetes_date >= "2020-01-01" ~ "y2020"),
                             levels = c("y2010", "y2015", "y2020")),
         date_range_2020 = relevel(date_range, ref = "y2020"))
```

# Calculate complicate dates

```{r}
complication_names <- c("mi", "stroke", "neuropathy", "esrd", "nephropathy", "retinopathy", "nafld", "dka", "u_ketone_first_positive")
for (c in complication_names){
  diagnosis_date <- paste(c, "date", sep = "_")
  joined_data <- joined_data %>%
    mutate("to_{c}" := case_when(!is.na(.[[diagnosis_date]]) ~ as.numeric(.[[diagnosis_date]] - diabetes_date, "days")/365,
                                 is.na(.[[diagnosis_date]]) & pandemic == "pre" ~ 
                                   as.numeric(mdy("12-31-2019") - diabetes_date, "days")/365,
                                 is.na(.[[diagnosis_date]]) & pandemic == "post" ~ 
                                   as.numeric(ymd(covid_censored_date) - diabetes_date, "days")/365,
                                 .default = NA) 
          )
}

joined_data <- joined_data %>% 
  mutate(to_deceased = case_when(!is.na(deceased_date) ~ as.numeric(deceased_date - diabetes_date, "days")/365,
                                 is.na(deceased_date) & pandemic == "pre" ~ 
                                   as.numeric(ymd(deceased_censor_date) - diabetes_date, "days")/365,
                                 is.na(deceased_date) & pandemic == "post" ~ 
                                   as.numeric(ymd(covid_censored_date) - diabetes_date, "days")/365,
                                 .default = NA))
```


# People with multiple HOMA measurements

```{r}
homa_path <- "/data/user/brianlu/projects/precision_diabetes/COVID19_clusters/exports/features_Ahlqvist_2024-06-28/homa_cpeptide_calc.csv"
homa <- read_csv(homa_path) %>% mutate(homa_cpeptide_date = mdy(homa_cpeptide_date))

covid_homa <- post_covid %>% 
  filter(covid19 == 1) %>% 
  select(DE_ID, covid19, covid19_date) %>% left_join(homa, by = join_by(DE_ID))

multiple_homa_tests <- covid_homa %>% group_by(DE_ID) %>% summarise(n = n()) %>% filter(n > 1) %>% pull(DE_ID)

covid_homa <- covid_homa %>% 
  filter(DE_ID %in% multiple_homa_tests) %>% 
  mutate(before_covid = case_when(homa_cpeptide_date <= covid19_date ~ "before",
                                  .default = "after"),
         to_covid19 = as.numeric(homa_cpeptide_date - covid19_date, "days")) %>%
  group_by(DE_ID, before_covid) %>%
  filter(to_covid19 == min(to_covid19)) # keep value closest to before and after covid

before <- covid_homa %>% group_by(DE_ID) %>%
  filter(n() > 1 & before_covid == "before")
after <- covid_homa %>% group_by(DE_ID) %>%
  filter(n() > 1 & before_covid == "after")  
  
t.test(before$homa_cpeptide_ir, after$homa_cpeptide_ir, paired = T)
```

# Import analysis functions

Source functions to generate summary tables.

```{r}
# import functions to summarize features into tables
source("/data/user/brianlu/projects/precision_diabetes/COVID19_clusters/functions_analysis.R")
```

Select features to summarize.

```{r}
# declare features
ahlqvist <- c("gad", "a1c", "bmi", "diabetes_age", "homa_cpeptide_b", "homa_cpeptide_ir")
demo <- c(# "median_income", 
          "deceased_age")
complications <- c("neuropathy", "retinopathy", "nephropathy", "esrd", "nafld", 
                   "cvd", # cumulative of mi and stroke
                   "mi", "stroke", 
                   "dka", "u_ketone_first_positive", "deceased",
                   "covid19")
glucose_homeostasis <- c("glucose", "insulin", "cpeptide", "homa_cpeptide_s")
liver <- c("albumin", "alt", "ast")
lipid <- c("cholesterol", "hdl", "ldl", "ldl_hdl", "non_hdl", "triglyceride", "triglyceride_hdl")
kidney <- c("creatinine", "gfr", "u_albumin", "u_creatinine")
prescription <- c("insulin_prescription")

# declare covariates
covariates <- c("race", "gender", "diabetes_age")

# declare list to store analysis results
analysis_list <- c()

# replace feature names from R variable naming conventions style to manuscript style
feature_rename <- function(df){
  df <- df %>% 
    mutate(Features = case_when(
      Features == "gad" ~ "GAD antibody",
      Features == "a1c" ~ "HbA1c",
      Features == "bmi" ~ "BMI",
      Features == "diabetes_age" ~ "Diagnosis age",
      Features == "homa_cpeptide_b" ~ "HOMA2-B",
      Features == "homa_cpeptide_ir" ~ "HOMA2-IR",
      Features == "gender" ~ "Gender",
      Features == "race" ~ "Race",
      Features == "median_income" ~ "Median income",
      Features == "deceased_age" ~ "Deceased age",
      Features == "neuropathy" ~ "Neuropathy",
      Features == "retinopathy" ~ "Retinopathy",
      Features == "nephropathy" ~ "Nephropathy",
      Features == "esrd" ~ "End-stage renal disease (ESRD)",
      Features == "nafld" ~ "MASLD",
      Features == "cvd" ~ "Cardiovascular disease (CVD)",
      Features == "mi" ~ "MI",
      Features == "stroke" ~ "Stroke",
      Features == "dka" ~ "Diabetic ketoacidosis (DKA)",
      Features == "u_ketone_first_positive" ~ "Urine ketones",
      Features == "deceased" ~ "Deceased",
      Features == "covid19" ~ "COVID19",
      Features == "glucose" ~ "Glucose",
      Features == "insulin" ~ "Insulin",
      Features == "cpeptide" ~ "C-peptide",
      Features == "homa_cpeptide_s" ~ "HOMA2-S",
      Features == "albumin" ~ "Albumin",
      Features == "alt" ~ "ALT",
      Features == "ast" ~ "AST",
      Features == "cholesterol" ~ "Cholesterol",
      Features == "hdl" ~ "HDL",
      Features == "ldl" ~ "LDL",
      Features == "ldl_hdl" ~ "LDL/HDL ratio",
      Features == "non_hdl" ~ "Non-HDL cholesterol",
      Features == "triglyceride" ~ "Triglycerides",
      Features == "triglyceride_hdl" ~ "Triglyceride/HDL ratio",
      Features == "creatinine" ~ "Creatinine",
      Features == "gfr" ~ "eGFR",
      Features == "u_albumin" ~ "Urine albumin",
      Features == "u_creatinine" ~ "Urine creatinine",
      Features == "insulin_prescription" ~ "Insulin prescription",
      .default = Features
    ))
  
  return(df)
}
```

Set reference levels for logistic regression.

```{r}
# relevel to set reference levels for regression
joined_data <- joined_data %>%
  mutate(race = factor(race, levels = c("White", "Black or African American", "Other")),
         gender = factor(gender, levels = c("Male", "Female")),
         pandemic = factor(pandemic, levels = c("pre", "post")),
         covid19_status = factor(case_when(!is.na(covid19_date) ~ "positive", 
                                           .default = "negative"), 
                                 levels = c("negative", "positive")),
         cluster = factor(cluster, levels = c("SAID", "SIDD", "SIRD", "MOD", "MARD")))

# joined_data$race_factor <- factor(joined_data$race)
# joined_data$race_factor <- relevel(joined_data$race_factor, ref = "White")
# 
# joined_data$gender_factor <- factor(joined_data$gender)
# joined_data$gender_factor <- relevel(joined_data$gender_factor, ref = "Male")
```


Set up export directory.

```{r}
# create export directory if doesn't exist
export_dir <- "/data/user/brianlu/projects/precision_diabetes/COVID19_clusters/exports"
export_dir <- file.path(export_dir, paste0("analysis_", Sys.Date()))
if (!dir.exists(export_dir)) {
  dir.create(export_dir, recursive = T)
}
```


# Table 1
Construct table 1 as tibble.

```{r}
# calculate stats for table
baseline <- get_feature_summary(df = joined_data, 
                                features = c("gender", "race", 
                                             "a1c", "bmi", "diabetes_age", 
                                             "deceased_age",
                                             "homa_cpeptide_ir", "glucose", "cpeptide", 
                                             "cholesterol", "ldl", "hdl", "triglyceride"
                                             ),
                                categories = c("pandemic"),
                                covariates = NA,
                                distribution_stat = "ci") %>% feature_rename()

# construct patient_char tibble
patient_char <- tibble(Period = c("N"), 
                       `2010-2019\n(Pre-pandemic)` = as.character(nrow(pre_covid)),
                       `2020-2024\n(Post-pandemic)` = as.character(nrow(post_covid)),
                       P = "")

# construct counts table to join with patient_char
counts_table <- baseline %>% 
  filter(grepl("count", Summary)) %>% # filter to count data only
  mutate(Value = str_replace(Value, "Black or African American", "AA"),
         Value = factor(Value, levels = c("Male", 
                                          "Female", 
                                          "AA", 
                                          "White", 
                                          "Other"))) %>%
  arrange(Value) %>% #rearrange order by factor
  rename(Period = Value, 
         `2010-2019\n(Pre-pandemic)` = pre, 
         `2020-2024\n(Post-pandemic)` = post, 
         P = `Pandemic p-value`) %>% # rename columns to match
  filter(Period != "Female") %>% # remove rows
  filter(Period != "Other") %>% # remove rows
  select(Period, `2010-2019\n(Pre-pandemic)`, `2020-2024\n(Post-pandemic)`, P) # select joining columns

# constructs continuous table to join with patient_char
cont_table <- baseline %>% 
  filter(!grepl("count", Summary)) %>% # filter to continuous data only
  mutate(Value = str_replace_all(Value, "Index|\\^", ""),
         Value = str_replace_all(Value, "Years", "years"),
         Period = case_when(Value != "" ~ paste0(Features, " [", Value, "]"),
                            .default = paste0(Features, Value))) %>% # attach units to feature
  rename(`2010-2019\n(Pre-pandemic)` = pre, 
         `2020-2024\n(Post-pandemic)` = post, 
         P = `Pandemic p-value`) %>% # rename columns to match
  select(Period, `2010-2019\n(Pre-pandemic)`, `2020-2024\n(Post-pandemic)`, P) # select joining columns

analysis_list[["patient_char"]] <- bind_rows(patient_char, counts_table, cont_table)

```

Add formatting and export as excel file.

```{r}
# create workbook
wb <- createWorkbook()

# add table 1
addWorksheet(wb, sheetName = "Table 1", gridLines = F)
writeData(wb, sheet = "Table 1", x = analysis_list[["patient_char"]], borders = "none", 
          headerStyle = createStyle(textDecoration = "Bold", 
                                    border = c("top", "bottom"), 
                                    halign = "center", 
                                    valign = "center"))
# bold first column
addStyle(wb, "Table 1", createStyle(textDecoration = "Bold", 
                                    halign = "left", 
                                    valign = "center"), 
         rows = 2:(nrow(analysis_list[["patient_char"]]) + 1), # skip header row
         cols = 1, gridExpand = T)

# center align sheet except first and last column
addStyle(wb, "Table 1", createStyle(textDecoration = NULL, 
                                     halign = "center", 
                                     valign = "center"), 
         rows = 1:nrow(analysis_list[["patient_char"]]) + 1, # skip header row
         cols = 2:ncol(analysis_list[["patient_char"]]), gridExpand = T) # except first column

# italicize p-value
addStyle(wb, "Table 1", createStyle(textDecoration = "italic", 
                                     halign = "center", 
                                     valign = "center", 
                                     fontSize = 9), 
         rows = 2:nrow(analysis_list[["patient_char"]]) + 1, # skip header row
         cols = ncol(analysis_list[["patient_char"]]), gridExpand = T)

# set column width
setColWidths(wb, "Table 1", cols = 1:ncol(analysis_list[["patient_char"]]), widths = "auto")

# export
saveWorkbook(wb, file.path(export_dir, "table_1.xlsx"), overwrite = TRUE)

```

# Figure 1

```{r}


cluster_colors = c("cornflowerblue", "powderblue", "darkseagreen", "lightsalmon", "plum")
dpi = 600

pie <- list()

# function to calculate numbers needed to generate pie chart
calc_pie <- function(df, group, label_order){
  percent <- prop.table(table(df[[group]]))
  pie_stats <- tibble(group = label_order,
                  percent = as.vector(percent[order(factor(names(percent), levels = label_order))]) * 100) %>%
    mutate(group = factor(group, levels = label_order)) %>%
    arrange(desc(group)) %>%
    mutate(ylabel = sprintf("%.0f", .$percent),
           ypos = cumsum(percent)- 0.5*percent)
  
  return(pie_stats)
}


# calculate numbers needed for pie chart
for (p in c("pre", "post")){
  pandemic_subset <- joined_data %>% 
    filter(pandemic == p) %>% 
    mutate(race = case_when(race == "Black or African American" ~ "AA", .default = race))
  
  pie[[paste(p, "overall", sep = "_")]] <- calc_pie(pandemic_subset, group = "cluster", label_order = order)
  # subset by gender
  for (g in c("Male", "Female")){
    pandemic_subset_gender <- pandemic_subset %>% filter(gender == g)
    
    pie[[paste(p, "overall", g, sep = "_")]] <- calc_pie(pandemic_subset_gender, group = "cluster", label_order = order)
  }
  # subset by gender
  for (r in c("White", "AA")){
    pandemic_subset_race <- pandemic_subset %>% filter(race == r)
    
    pie[[paste(p, "overall", r, sep = "_")]] <- calc_pie(pandemic_subset_race, group = "cluster", label_order = order)
  } 
  
  
  pie[[paste(p, "gender", sep = "_")]] <- calc_pie(pandemic_subset, group = "gender", label_order = c("Male", "Female"))
  pie[[paste(p, "race", sep = "_")]] <- calc_pie(pandemic_subset, group = "race", label_order = c("White", "AA", "Other"))

}

for (n in names(pie)){

  # generate base pie chart
  pie_graph <- ggplot(pie[[n]], aes(x = "", y = percent, fill = group)) +
    geom_bar(stat="identity", width = 1, show.legend = T) +
    coord_polar("y", start = pi/2)

  # add colors
  if (nrow(pie[[n]]) == 5){
    pie_graph <- pie_graph + scale_fill_manual(values = cluster_colors, limits = order)
  } else if (nrow(pie[[n]]) == 3){
    pie_graph <- pie_graph + scale_fill_brewer(palette = "Accent")
  } else if (nrow(pie[[n]]) == 2){
    pie_graph <- pie_graph + scale_fill_brewer(palette = "Set2")
  }
  
  # add percent labels and theme
  pie_graph <- pie_graph + geom_text(aes(x = 1.1, y = ypos, label = paste0(ylabel, "%")), color = "black", size = 8*0.5) +
      labs(fill = "") + 
      theme_classic() + 
      theme_void()

  # add legend if post pandemic overall
  if (n == "post_overall"){
    pie_graph <- pie_graph + theme(#plot.title = element_text(hjust = 0.5, size = 8),
      legend.text=element_text(size = 8),
      legend.justification = "top")
    
    # save larger size to accommodate for legend
    ggsave(file.path(export_dir, paste0("fig1_", n, ".tiff")), plot = pie_graph, width = 6.5, height = 4.5, units = "cm", dpi = dpi)

  } else { # remove legend for all others
    pie_graph <- pie_graph + theme(legend.position = "none")
    
    # save
    ggsave(file.path(export_dir, paste0("fig1_", n, ".tiff")), plot = pie_graph, width = 4.5, height = 4.5, units = "cm", dpi = dpi)
  }
  

  
}


```

## Gender comparision

```{r}

analysis_list[["uni_pre_pandemic_gender"]] <- get_feature_summary(df = joined_data %>% filter(pandemic == "pre"), 
                                                                  features = "gender", 
                                                                  categories = "cluster") %>%
  feature_rename()


analysis_list[["uni_post_pandemic_gender"]] <- get_feature_summary(df = joined_data %>% filter(pandemic == "post"), 
                                                                   features = "gender", 
                                                                   categories = "cluster") %>%
  feature_rename()


analysis_list[["uni_male"]] <- get_feature_summary(df = joined_data %>% filter(gender == "Male"), 
                                                                  features = "pandemic", 
                                                                  categories = "cluster") %>%
  feature_rename()

analysis_list[["uni_female"]] <- get_feature_summary(df = joined_data %>% filter(gender == "Female"), 
                                                                  features = "pandemic", 
                                                                  categories = "cluster") %>%
  feature_rename()
```

## Race comparison
Due to small numbers in other racial categories, only White and Black/African American patients were analyzed. 

```{r}
bw <- joined_data %>% filter(race != "Other") %>% mutate(race = factor(race, levels = c("White", "Black or African American")))

analysis_list[["uni_pre_pandemic_race"]] <- get_feature_summary(df = bw %>% filter(pandemic == "pre"), 
                                                                  features = "race", 
                                                                  categories = "cluster") %>%
  feature_rename()


analysis_list[["uni_post_pandemic_race"]] <- get_feature_summary(df = bw %>% filter(pandemic == "post"), 
                                                                   features = "race", 
                                                                   categories = "cluster") %>%
  feature_rename()


analysis_list[["uni_white"]] <- get_feature_summary(df = bw %>% filter(race == "White"), 
                                                                  features = "pandemic", 
                                                                  categories = "cluster") %>%
  feature_rename()

analysis_list[["uni_AA"]] <- get_feature_summary(df = bw %>% filter(race == "Black or African American"), 
                                                                  features = "pandemic", 
                                                                  categories = "cluster") %>%
  feature_rename()
```

# Figure 2

```{r}

propci <- function(x, n, conf_level = 0.95){
  
  # calculate critical value of given confidence level
  alpha <- 1 - conf_level
  alpha_crit <- 1-alpha/2
  crit <- qnorm(alpha_crit)
  
  # calculate proportion
  p <- x/n
  
  # calculate standard error of proportion
  se <- sqrt((p * (1 - p)) / n)
  
  # calculate margin of error
  me <- se * crit
  
  # calcualte upper and lower limits of confidence interval
  upper <- p + me
  lower <- p - me
  
  # clip ci to max 1 and min 0
  if (upper > 1){
    upper <- 1
  }
  
  if (lower < 0){
    lower <- 0
  }
  
  # return ci
  return(c(lower, upper))
}


joined_data <- joined_data %>% mutate(year = year(diabetes_date))

for (c in order){
  # calculate proportions
  prop <- joined_data %>% 
    group_by(year) %>% 
    reframe(cluster = sum(cluster == c), 
            n = n(), 
            proportion = cluster/n, 
            ci_lower = propci(cluster, n)[1], 
            ci_upper = propci(cluster, n)[2])
  
  analysis_list[[paste(c, "proportion", sep = "_")]] <- prop
  
  prop_graph <- ggplot(prop, mapping = aes(x = year, y = proportion)) +
    geom_point(size = 1, color = "darkgrey") +
    geom_errorbar(aes(x = year, ymin = ci_lower, ymax = ci_upper),  width = 0.5, color = "darkgrey") +
    geom_smooth(se = F, color = "red") +
    geom_vline(xintercept = 2020, linetype = "dashed", color = "blue") + 
    scale_x_continuous(labels = c("'10", "'15", "'20"), breaks = c(2010, 2015, 2020)) +
    scale_y_continuous(limits = c(0, 0.8)) + 
    theme_classic() + 
    labs(x = "Year", y = "Proportion") + 
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5),
          plot.margin = margin(0, 10, 0, 10)) # increase margin, clockwise from top, right, bottom, left
  
  # save
  ggsave(file.path(export_dir, paste0("fig2_", c, "_trend.tiff")), plot = prop_graph, width = 5, height = 4.5, units = "cm", dpi = dpi)
}




```

## Prop test

```{r}


prop_test_results <- joined_data %>% 
  # build count table
  group_by(cluster, pandemic) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = pandemic, values_from = n) %>% 
  ungroup()
  
prop_p <- c()
for (c in order){
  p <- prop.test(c(prop_test_results %>% filter(cluster == c) %>% pull(pre),
                   prop_test_results %>% filter(cluster == c) %>% pull(post)),
                 c(nrow(pre_covid), nrow(post_covid)))
  prop_p <- append(prop_p, p$p.value)
}

analysis_list[["uni_cluster_proportions"]] <- prop_test_results %>% 
  mutate(z_test_p_value = prop_p)


```


## Per cluster comparison

```{r}
clusters <- c("SAID", "SIDD", "SIRD", "MOD", "MARD")

for (c in clusters)
  analysis_list[[paste("uni", c, sep = "_")]] <- get_feature_summary(df = joined_data %>% filter(cluster == c), 
                                                                     features = c("race", "gender",
                                                                                  ahlqvist, 
                                                                                  #"deceased_age", # too few in SAID
                                                                                  glucose_homeostasis,
                                                                                  lipid,
                                                                                  kidney),
                                                                     categories = c("pandemic"),
                                                                     covariates = NA) %>% feature_rename()
```


# Figure 3

```{r}

# within SIDD, proportion of each race
for (r in c("White", "Black or African American")){
  # calculate proportions
  prop <- joined_data %>% 
    filter(cluster == "SIDD") %>%
    group_by(year) %>% 
    reframe(race = sum(race == r), 
            n = n(), 
            proportion = race/n, 
            ci_lower = propci(race, n)[1], 
            ci_upper = propci(race, n)[2])
  
  analysis_list[[paste(r, "SIDD_proportion", sep = "_")]] <- prop
  
  prop_graph <- ggplot(prop, mapping = aes(x = year, y = proportion)) +
    geom_point(size = 1, color = "darkgrey") +
    geom_errorbar(aes(x = year, ymin = ci_lower, ymax = ci_upper),  width = 0.5, color = "darkgrey") +
    geom_smooth(se = F, color = "red") +
    geom_vline(xintercept = 2020, linetype = "dashed", color = "blue") + 
    scale_x_continuous(labels = c("'10", "'15", "'20"), breaks = c(2010, 2015, 2020)) +
    theme_classic() + 
    labs(x = "Year", y = "Proportion") + 
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5),
          plot.margin = margin(0, 10, 0, 10)) # increase margin, clockwise from top, right, bottom, left
  
  # save
  ggsave(file.path(export_dir, paste("fig3", "SIDD", r, "trend.tiff", sep = "_")), plot = prop_graph, width = 5, height = 4.5, units = "cm", dpi = dpi)
}


# Proportion of each race that is SIDD
for (r in c("White", "Black or African American")){
  # calculate proportions
  prop <- joined_data %>% 
    filter(race == r) %>%
    group_by(year) %>% 
    reframe(cluster = sum(cluster == "SIDD"), 
            n = n(), 
            proportion = cluster/n, 
            ci_lower = propci(cluster, n)[1], 
            ci_upper = propci(cluster, n)[2])
  
  analysis_list[[paste(r, "SIDD_proportion", sep = "_")]] <- prop
  
  prop_graph <- ggplot(prop, mapping = aes(x = year, y = proportion)) +
    geom_point(size = 1, color = "darkgrey") +
    geom_errorbar(aes(x = year, ymin = ci_lower, ymax = ci_upper),  width = 0.5, color = "darkgrey") +
    geom_smooth(se = F, color = "red") +
    geom_vline(xintercept = 2020, linetype = "dashed", color = "blue") + 
    scale_x_continuous(labels = c("'10", "'15", "'20"), breaks = c(2010, 2015, 2020)) +
    scale_y_continuous(limits = c(0, 1)) + 
    theme_classic() + 
    labs(x = "Year", y = "Proportion") + 
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5),
          plot.margin = margin(0, 10, 0, 10)) # increase margin, clockwise from top, right, bottom, left
  
  # save
  ggsave(file.path(export_dir, paste("fig3", r, "SIDD_trend.tiff", sep = "_")), plot = prop_graph, width = 5, height = 4.5, units = "cm", dpi = dpi)
}

```


Proportion test

```{r}
# for proportion of SIDD in each race
prop_test_results <- joined_data %>% 
  # build count table
  filter(cluster == "SIDD") %>%
  group_by(race, pandemic) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = pandemic, values_from = n) %>% 
  ungroup()

prop_p <- c()
for (r in prop_test_results$race){
  p <- prop.test(c(prop_test_results %>% filter(race == r) %>% pull(pre),
                   prop_test_results %>% filter(race == r) %>% pull(post)),
                 c(nrow(pre_covid %>% filter(race == r)), 
                   nrow(post_covid %>% filter(race == r))))
  prop_p <- append(prop_p, p$p.value)
}

analysis_list[["uni_SIDD_race_proportions"]] <- prop_test_results %>% 
  mutate(z_test_p_value = prop_p)

# within SIDD, proportion of each race
prop_test_results <- joined_data %>% 
  # build count table
  filter(cluster == "SIDD") %>%
  group_by(race, pandemic) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = pandemic, values_from = n) %>% 
  ungroup()

prop_p <- c()
for (r in prop_test_results$race){
  p <- prop.test(c(prop_test_results %>% filter(race == r) %>% pull(pre),
                   prop_test_results %>% filter(race == r) %>% pull(post)),
                 c(nrow(pre_covid %>% filter(cluster == "SIDD")), 
                   nrow(post_covid %>% filter(cluster == "SIDD"))))
  prop_p <- append(prop_p, p$p.value)
}

analysis_list[["uni_SIDD_race_proportions"]] <- prop_test_results %>% 
  mutate(z_test_p_value = prop_p)
```




```{r}

box_theme <- theme(axis.title.x = element_text(size = 8, color = "black"),
                    axis.text.x = element_text(size = 8, color = "black"),
                    axis.title.y = element_text(size = 8, color = "black"),
                    axis.text.y = element_text(size = 8, color = "black"),
                    axis.line = element_line(linewidth = 0.25, color = "black"),
                    axis.ticks = element_line(linewidth = 0.25, color = "black"),
                    plot.title = element_text(hjust = 0.5),
                    legend.text=element_text(size = 8))

age <- ggplot(data = joined_data, mapping = aes(x = pandemic, y = diabetes_age)) + 
  geom_boxplot() +
  theme_classic() + 
  scale_x_discrete(labels = c("2010-2019", "2020-2024")) +
  labs(x = "Year", y = "Diagnosis age (years)") + 
  box_theme

ggsave(file.path(export_dir, paste0("box_age.tiff")), plot = age, width = 5, height = 4.5, units = "cm", dpi = dpi)

bmi <- ggplot(data = joined_data, mapping = aes(x = pandemic, y = bmi)) + 
  geom_boxplot() +
  theme_classic() + 
  scale_x_discrete(labels = c("2010-2019", "2020-2024")) +
  labs(x = "Year", y = bquote("BMI " (kg/m^2))) + 
  box_theme

ggsave(file.path(export_dir, paste0("box_bmi.tiff")), plot = bmi, width = 5, height = 4.5, units = "cm", dpi = dpi)

deceased_age <- ggplot(data = joined_data, mapping = aes(x = pandemic, y = deceased_age)) + 
  geom_boxplot() +
  theme_classic() + 
  scale_x_discrete(labels = c("2010-2019", "2020-2024")) +
  labs(x = "Year", y = "Deceased age (years)") + 
  box_theme

ggsave(file.path(export_dir, paste0("box_deceased_age.tiff")), plot = deceased_age, width = 5, height = 4.5, units = "cm", dpi = dpi)

ir <- ggplot(data = joined_data, mapping = aes(x = pandemic, y = homa_cpeptide_ir)) + 
  geom_boxplot() +
  theme_classic() + 
  scale_x_discrete(labels = c("2010-2019", "2020-2024")) +
  labs(x = "Year", y = "HOMA2-IR") + 
  box_theme

ggsave(file.path(export_dir, paste0("box_homa2ir.tiff")), plot = ir, width = 5, height = 4.5, units = "cm", dpi = dpi)


```


# Inpatient encounters

```{r}
get_feature_summary(df = joined_data, features = "inpatient", categories = "pandemic")

joined_data %>% finalfit::finalfit("Surv(to_inpatient, inpatient)", c("race", "gender", "diabetes_age", "pandemic"))

inpatient <- joined_data %>% filter(inpatient == 1)


```

```{r}
inpatient %>% finalfit::finalfit("Surv(to_deceased, deceased)", c("race", "gender", "diabetes_age", "pandemic"))
```


Export results for graphing.

```{r}
# create export directory if doesn't exist
export_dir <- "/data/user/brianlu/projects/precision_diabetes/COVID19_clusters/exports"
export_dir <- file.path(export_dir, paste0("analysis_", Sys.Date()))
if (!dir.exists(export_dir)) {
  dir.create(export_dir, recursive = T)
}

# source data
write_csv(joined_data, file.path(export_dir, "joined_data.csv"))

# analysis
#saveWorkbook(wb, file.path(export_dir, "characteristic_table.xlsx"), overwrite = TRUE)

for (n in names(analysis_list)){
  write_csv(analysis_list[[n]], file.path(export_dir, paste0(n, ".csv")))
}
```

# Session Information

```{r}
sessionInfo()
```


fig1: pie chart for pre vs post, overall, male vs female, race
fig2: trend line of proportion per year per cluster
z test for pre vs post

Per cluster and overall homa-ir, age, a1c, bmi

refernces for covid and beta cells and insulin resistance

outcome from hospital
inpatient encounter, time to death, total inpatient encounter, per cluster



```{r}
# test population level SIRD
prop.test(c(sum(pre_covid$SIRD == 1), sum(post_covid$SIRD == 1)),
          c(nrow(pre_covid), nrow(post_covid)))

# test population level MARD
prop.test(c(sum(pre_covid$MARD == 1), sum(post_covid$MARD == 1)),
          c(nrow(pre_covid), nrow(post_covid)))

# 
t.test(pre_covid$diabetes_age, post_covid$diabetes_age)
t.test(pre_covid$homa_cpeptide_ir, post_covid$homa_cpeptide_ir)

pre_sird <- pre_covid %>% filter(cluster == "SIRD")
post_sird <- post_covid %>% filter(cluster == "SIRD")

t.test(pre_sird$diabetes_age, post_sird$diabetes_age)
```


```{r}
joined_data <- joined_data %>% 
  mutate(severe = factor(case_when(cluster == "SIDD" | cluster == "SIRD" ~ "severe", 
                                   cluster == "SAID" ~ "autoimmune",
                                   cluster == "MARD" | cluster == "MOD" ~ "mild"),
                         levels = c("mild", "severe", "autoimmune")))

joined_data <- joined_data %>% 
  mutate(severe = factor(case_when(cluster == "SIDD" | cluster == "SIRD" ~ "severe", 
                                   cluster == "MOD" | cluster == "MARD" ~ "mild"),
                         level = c("mild", "severe")),
         SIDD_factor = factor(case_when(cluster == "SIDD" ~ "SIDD", .default = "Others"), 
                              levels = c("Others", "SIDD")))

joined_data <- joined_data %>% 
  mutate(severe = factor(case_when(cluster == "SIDD" | cluster == "SIRD" ~ "severe", 
                                   cluster == "MOD" | cluster == "MARD"| cluster == "SAID"~ "mild"),
                         level = c("mild", "severe")),
         SIDD_factor = factor(case_when(cluster == "SIDD" ~ "SIDD", .default = "Others"), 
                              levels = c("Others", "SIDD")))

joined_data <- joined_data %>% 
  mutate(b_deficient = factor(case_when(cluster == "SIDD" | cluster == "SAID" ~ "deficient", 
                                 .default = "not_deficient"), levels = c("not_deficient", "deficient")))

joined_data <- joined_data %>%
  mutate(year = year(diabetes_date),
         year_increment = case_when(date_range == "y2010" ~ 0,
                                    date_range == "y2015" ~ 1,
                                    date_range == "y2020" ~ 2),
         year_2 = floor(year(diabetes_date)/2-1005),
         age_factor = floor(diabetes_age/10))

joined_data %>% group_by(date_range) %>% summarise(SAID = sum(SAID == 1)/n(), 
                                             SIDD = sum(SIDD == 1)/n(),
                                             SIRD = sum(SIRD == 1)/n(),
                                             MOD = sum(MOD == 1)/n(),
                                             MARD = sum(MARD == 1)/n()) %>%
  pivot_longer(cols = SAID:MARD, names_to = "cluster", values_to = "Proportion") %>% 
  mutate(cluster = factor(cluster, levels = c("SAID", "SIDD", "SIRD", "MOD", "MARD", "severe"))) %>% 
  ggplot(., mapping = aes(x = date_range, y = Proportion)) + geom_bar(stat = "identity") + theme_classic() +
    facet_wrap(vars(cluster))


joined_data %>% group_by(year_2) %>% summarise(SAID = sum(SAID == 1)/n(), 
                                             SIDD = sum(SIDD == 1)/n(),
                                             SIRD = sum(SIRD == 1)/n(),
                                             MOD = sum(MOD == 1)/n(),
                                             MARD = sum(MARD == 1)/n(),
                                             severe = sum(severe == "severe")/n()) %>%
  pivot_longer(cols = SAID:severe, names_to = "cluster", values_to = "Proportion") %>% 
  mutate(cluster = factor(cluster, levels = c("SAID", "SIDD", "SIRD", "MOD", "MARD", "severe"))) %>% 
  ggplot(., mapping = aes(x = year_2, y = Proportion)) + geom_smooth() + geom_point() + theme_classic() + geom_vline(xintercept = 5, linetype = "dashed") +
    facet_wrap(vars(cluster))

joined_data %>% group_by(year) %>% summarise(SAID = sum(SAID == 1)/n(), 
                                             SIDD = sum(SIDD == 1)/n(),
                                             SIRD = sum(SIRD == 1)/n(),
                                             MOD = sum(MOD == 1)/n(),
                                             MARD = sum(MARD == 1)/n(),
                                             Severe = sum(severe == "severe")/n(),
                                             Mild = sum(severe == "mild")/n()) %>%
  pivot_longer(cols = SAID:Mild, names_to = "cluster", values_to = "Proportion") %>% 
  mutate(cluster = factor(cluster, levels = c("SAID", "SIDD", "SIRD", "MOD", "MARD", "Severe", "Mild"))) %>% 
  ggplot(., mapping = aes(x = year, y = Proportion)) + geom_smooth() + geom_point() + theme_classic() + geom_vline(xintercept = 2020, linetype = "dashed") +
    facet_wrap(vars(cluster))


joined_data %>% group_by(year, covid19_status) %>% summarise(
                                             SAID = sum(SAID == 1)/n(), 
                                             SIDD = sum(SIDD == 1)/n(),
                                             SIRD = sum(SIRD == 1)/n(),
                                             MOD = sum(MOD == 1)/n(),
                                             MARD = sum(MARD == 1)/n(),
                                             severe = sum(severe == "severe")/n()) %>%
  pivot_longer(cols = SAID:severe, names_to = "cluster", values_to = "Proportion") %>% 
  mutate(cluster = factor(cluster, levels = c("SAID", "SIDD", "SIRD", "MOD", "MARD", "severe"))) %>% 
  ggplot(., mapping = aes(x = year, y = Proportion, color = covid19_status)) + geom_smooth() + geom_point() + theme_classic() + geom_vline(xintercept = 2020, linetype = "dashed") +
    facet_wrap(vars(cluster))

joined_data %>% group_by(year) %>% summarise(SIRD = sum(SIRD == 1)/n(),
                                             MARD = sum(MARD == 1)/n(),
                                             Others = sum(cluster != "SIRD" & cluster != "MARD")/n()) %>%
  pivot_longer(cols = SIRD:Others, names_to = "cluster", values_to = "Proportion") %>% 
  mutate(cluster = factor(cluster, levels = c("SIRD", "MARD", "Others"))) %>% 
  ggplot(., mapping = aes(x = year, y = Proportion)) + geom_smooth() + geom_point() + theme_classic() + geom_vline(xintercept = 2020, linetype = "dashed") +
    facet_wrap(vars(cluster))


joined_data %>% select(diabetes_date, a1c, bmi, diabetes_age, homa_cpeptide_b, homa_cpeptide_ir) %>%
  pivot_longer(cols = a1c:homa_cpeptide_ir, names_to = "feature", values_to = "value") %>%
  ggplot(., mapping = aes(x = diabetes_date, y = value)) + 
  geom_point(alpha = 0.25, size = 0.25) + 
  geom_smooth() + 
  facet_wrap(vars(feature), scale = "free") + 
  theme_classic()

joined_data %>% select(severe, diabetes_date, a1c, bmi, diabetes_age, homa_cpeptide_b, homa_cpeptide_ir) %>%
  pivot_longer(cols = a1c:homa_cpeptide_ir, names_to = "feature", values_to = "value") %>%
  ggplot(., mapping = aes(x = diabetes_date, y = value, color = severe)) + 
  geom_point(alpha = 0.25, size = 0.25) + 
  geom_smooth() + 
  facet_wrap(vars(feature), scale = "free") + 
  theme_classic()

ggplot(joined_data, mapping = aes(x = diabetes_date, y = homa_cpeptide_ir, color = cluster)) +
  geom_point(alpha = 0.25, size = 0.25) +
  geom_smooth() + 
  coord_cartesian(ylim = c(0, 10)) + 
  theme_classic()


joined_data %>% group_by(date_range) %>% summarise(
    SAID = sum(SAID == 1)/n(), 
    SIDD = sum(SIDD == 1)/n(),
    SIRD = sum(SIRD == 1)/n(),
    MOD = sum(MOD == 1)/n(),
    MARD = sum(MARD == 1)/n(),
    severe = sum(severe == "severe")/n()) %>%
  pivot_longer(cols = SAID:severe, names_to = "cluster", values_to = "Proportions") %>%
  mutate(cluster = factor(cluster, levels = c("SAID", "SIDD", "SIRD", "MOD", "MARD", "severe"))) %>%
  ggplot(., mapping = aes(x = date_range, y = Proportions)) + geom_bar(stat = "identity") + facet_wrap(vars(cluster))


```


```{r}
chisq.test(table(joined_data$cluster, joined_data$date_range))

summary(glm(b_deficient ~ race + gender + diabetes_age + date_range + covid19_status, data = joined_data, family = "binomial"))

summary(lm(homa_cpeptide_b ~ race + gender + diabetes_age + date_range + homa_cpeptide_ir + SIDD_factor * covid19_status,
           data = joined_data))


```

# time to covid
```{r}
joined_data <- joined_data %>% 
  mutate(gad_covid = case_when(gad_date >= covid19_date ~ 1, .default = 0),
         a1c_covid = case_when(a1c_date >= covid19_date ~ 1, .default = 0),
         bmi_covid = case_when(bmi_date >= covid19_date ~ 1, .default = 0),
         homa_covid = case_when(homa_cpeptide_date >= covid19_date ~ 1, .default = 0),
         diabetes_covid = case_when(diabetes_date >= covid19_date ~ 1, .default = 0))



pre_covid_DE_ID <- joined_data %>% 
  filter(covid19 == 1 & diabetes_covid == 0 & gad_covid == 0 & a1c_covid == 0 & bmi_covid == 0 & homa_covid == 0) %>% 
  pull(DE_ID)

covid_history <- post_covid %>% filter(DE_ID %in% pre_covid_DE_ID) %>% filter(covid19_code != "U07.1") %>% pull(DE_ID)

# how do the 6 patients in covid_history line up with test dates?

joined_data <- joined_data %>%
  mutate(after_covid = case_when(diabetes_date >= covid19_date ~ 1,
                                 gad_date >= covid19_date ~ 1,
                                 a1c_date >= covid19_date ~ 1,
                                 bmi_date >= covid19_date ~ 1,
                                 homa_cpeptide_date >= covid19_date ~ 1,
                                 # DE_ID %in% covid_history ~ 1,
                                 .default = 0))

```


# Complications

```{r}
complication_names <- c("mi", "stroke", "neuropathy", "esrd", "nephropathy", "retinopathy", "nafld", "dka", "u_ketone_first_positive")
for (c in complication_names){
  diagnosis_date <- paste(c, "date", sep = "_")
  joined_data <- joined_data %>%
    mutate("to_{c}" := case_when(!is.na(.[[diagnosis_date]]) ~ as.numeric(.[[diagnosis_date]] - diabetes_date, "days")/365,
                                 is.na(.[[diagnosis_date]]) & pandemic == "pre" ~ 
                                   as.numeric(mdy("12-31-2019") - diabetes_date, "days")/365,
                                 is.na(.[[diagnosis_date]]) & pandemic == "post" ~ 
                                   as.numeric(ymd(covid_censored_date) - diabetes_date, "days")/365,
                                 .default = NA) 
          )
}


analysis <- list()
for (i in c(complication_names)){
  analysis[[i]] <- joined_data %>% finalfit::finalfit(paste0("Surv(to_", i, ", ", i, ")"), 
                                c(covariates, "date_range", "cluster"), pval = TRUE)
}
```


```{r}
joined_data %>% finalfit::surv_plot("Surv(to_dka, dka)", "pandemic", fun = "event")
```



















# Overall risks of complications
Analyze risks of complications in the entire cohort, in the AA patients only, and in the White patients only.

```{r}
# filter post_covid by race
aa <- post_covid %>% filter(race == "Black or African American")
white <- post_covid %>% filter(race == "White")

# risks of MI by race
for (comp in complications){
  analysis_list[[paste("multi_overall", comp, sep = "_")]] <- logit_table(post_covid, comp, covariates) %>% 
    arrange(factor(Value, levels = c("White", "Black or African American", "Other",
                                     "Male", "Female")))
  analysis_list[[paste("multi_aa", comp, sep = "_")]] <- logit_table(aa, comp, c("gender", "diabetes_age"))
  analysis_list[[paste("multi_white", comp, sep = "_")]] <- logit_table(white, comp, c("gender", "diabetes_age"))
  
}

```

Analyze association between MI and potential risk factors.

```{r}
# association between MI and potential risk factors
mi_features <- c("a1c",
                 "homa_cpeptide_b_25",
                 "homa_cpeptide_ir",
                 "bmi_5",
                 "cholesterol_50",
                 "ldl_hdl",
                 "non_hdl_50")

for (m in mi_features){
  analysis_list[[paste("multi_overall_mi", m, sep = "_")]] <- logit_table(post_covid, "mi", c(m, covariates)) %>% 
    arrange(factor(Value, levels = c("White", "Black or African American", "Other",
                                     "Male", "Female")))
  analysis_list[[paste("multi_aa_mi", m, sep = "_")]] <- logit_table(aa, "mi", c(m, "gender", "diabetes_age"))
  analysis_list[[paste("multi_white_mi", m, sep = "_")]] <- logit_table(white, "mi", c(m, "gender", "diabetes_age"))
}

```

MI is associated with HbA1c and HOMA2-B in AA but not Whites. AA is known to have higher risks of nephropathy, which is associated with increased risks of MI. 

```{r}
neph_adjust <- c("a1c", "homa_cpeptide_b_25")

for (n in neph_adjust){
  analysis_list[[paste("multi_overall_mi_neph", n, sep = "_")]] <- logit_table(post_covid, "mi", c(n, covariates, "nephropathy")) %>% 
    arrange(factor(Value, levels = c("White", "Black or African American", "Other",
                                     "Male", "Female")))
  analysis_list[[paste("multi_aa_mi_neph", n, sep = "_")]] <- logit_table(aa, "mi", c(n, "gender", "diabetes_age", "nephropathy"))
  analysis_list[[paste("multi_white_mi_neph", n, sep = "_")]] <- logit_table(white, "mi", c(n, "gender", "diabetes_age", "nephropathy"))
  
  # interaction between race and a1c or homa2b
  formula <- reformulate(c("gender_factor", paste("race_factor", n, sep = "*")), response = "mi")
  model <- glm(formula, data = post_covid, family = "binomial")
  conf <- exp(confint.default(model)) %>% as_tibble(rownames = "term")
  coefs <- exp(coef(model)) %>% as_tibble(rownames = "term") %>% rename(OR = value)
  analysis_list[[paste("multi_interaction_mi", n, sep = "_")]] <- summary(model)$coefficient %>% 
    as_tibble(rownames = "term") %>%
    left_join(coefs, by = join_by(term)) %>%
    left_join(conf, by = join_by(term))
}

```

# Insulin vs c-peptide HOMA2 estimates
Construct new tibble to build linear regression models.

```{r}
homa <- post_covid %>% 
  filter(!is.na(homa_insulin_b) & !is.na(homa_cpeptide_b) & insulin_date == cpeptide_date) %>%
  #filter(insulin_date < insulin_order_date | is.na(insulin_order_date)) %>% # value must be prior to insulin prescription
  filter(race != "Other") %>%
  mutate(count = "1",
         homab_dif = homa_cpeptide_b - homa_insulin_b, 
         homair_dif = homa_cpeptide_ir - homa_insulin_ir ) # add to summarize by count later

ic_b <- homa %>% 
  select(DE_ID, race, gender, diabetes_age, homa_insulin_b, homa_cpeptide_b) %>%
  pivot_longer(cols = c(homa_insulin_b, homa_cpeptide_b), names_to = "method", values_to = "homa2b") %>%
  mutate(DE_ID = factor(DE_ID),
         method = ifelse(.$method == "homa_insulin_b", "insulin", "cpeptide"))

ic_ir <- homa %>%
  select(DE_ID, race, gender, diabetes_age, homa_insulin_ir, homa_cpeptide_ir) %>%
  pivot_longer(cols = c(homa_insulin_ir, homa_cpeptide_ir), names_to = "method", values_to = "homa2ir") %>%
  mutate(DE_ID = factor(DE_ID),
         method = ifelse(.$method == "homa_insulin_ir", "insulin", "cpeptide"))

homa_ic <- full_join(ic_b, ic_ir, by = join_by(DE_ID, race, gender, diabetes_age, method))
```

Summarize and run paired t tests. The difference between HOMA2-B estimated from c-peptide and insulin are not significant in White patients, but the difference between HOMA2-B estimated from c-peptide and insulin are significant in Black/African American patients.

```{r}
b_summary <- get_feature_summary(homa,
                                 c("homa_insulin_b", "homa_cpeptide_b"),
                                 "race",
                                 c("gender", "diabetes_age"),
                                 distribution_stat = "ci") %>%
  mutate(`Overall paired p-value` = t.test(homa2b ~ method, data = homa_ic, paired = T)$p.value,
         `White paired p-value` = t.test(homa2b ~ method, data = homa_ic %>% filter(race == "White"), paired = T)$p.value,
         `AA paired p-value` = t.test(homa2b ~ method, data = homa_ic %>% filter(race == "Black or African American"), paired = T)$p.value)

ir_summary <- get_feature_summary(homa,
                                 c("homa_insulin_ir", "homa_cpeptide_ir"),
                                 "race",
                                 c("gender", "diabetes_age"),
                                 distribution_stat = "ci",
                                 digits = 2) %>%
  mutate(`Overall paired p-value` = t.test(homa2b ~ method, data = homa_ic, paired = T)$p.value,
         `White paired p-value` = t.test(homa2ir ~ method, data = homa_ic %>% filter(race == "White"), paired = T)$p.value,
         `AA paired p-value` = t.test(homa2ir ~ method, data = homa_ic %>% filter(race == "Black or African American"), paired = T)$p.value)

count_summary <- get_feature_summary(homa, "count", "race")

analysis_list$uni_homa_race <- full_join(b_summary, ir_summary) %>% full_join(count_summary)
```

Differences within AA or White.

```{r}
# AA homa2-b
homa2b_for <- reformulate(termlabels = c(covariates[covariates != "race"], "method", "(1 | DE_ID)"), "homa2b")
mod <- lmer(homa2b_for, data = homa_ic %>% filter(race == "Black or African American"))
analysis_list$multi_homa2b_aa_lme <- anova(mod, type = "II", ddf = "Satterthwaite") %>% as_tibble(rownames = "terms")

# AA homa2-ir
homa2ir_for <- reformulate(termlabels = c(covariates[covariates != "race"], "method", "(1 | DE_ID)"), "homa2ir")
mod <- lmer(homa2ir_for, data = homa_ic %>% filter(race == "Black or African American"))
analysis_list$multi_homa2ir_aa_lme <- anova(mod, type = "II", ddf = "Satterthwaite") %>% as_tibble(rownames = "terms")

# White homa2-b
homa2b_for <- reformulate(termlabels = c(covariates[covariates != "race"], "method", "(1 | DE_ID)"), "homa2b")
mod <- lmer(homa2b_for, data = homa_ic %>% filter(race == "White"))
analysis_list$multi_homa2b_white_lme <- anova(mod, type = "II", ddf = "Satterthwaite") %>% as_tibble(rownames = "terms")

# White homa2-ir
homa2ir_for <- reformulate(termlabels = c(covariates[covariates != "race"], "method", "(1 | DE_ID)"), "homa2ir")
mod <- lmer(homa2ir_for, data = homa_ic %>% filter(race == "White"))
analysis_list$multi_homa2ir_white_lme <- anova(mod, type = "II", ddf = "Satterthwaite") %>% as_tibble(rownames = "terms")
```

Fit linear mixed-effects model to test for difference in race.

```{r}
# homa2-b
homa2b_for <- reformulate(termlabels = c(covariates, "method", "method:race", "(1 | DE_ID)"), "homa2b")
mod <- lmer(homa2b_for, data = homa_ic)
analysis_list$multi_homa2b_race_lme <- anova(mod, type = "II", ddf = "Satterthwaite") %>% as_tibble(rownames = "terms")

# homa2-ir
homa2ir_for <- reformulate(termlabels = c(covariates, "method", "method:race", "(1 | DE_ID)"), "homa2ir")
mod <- lmer(homa2ir_for, data = homa_ic)
analysis_list$multi_homa2ir_race_lme <- anova(mod, type = "II", ddf = "Satterthwaite") %>% as_tibble(rownames = "terms")
```

# Cluster characteristics

```{r}
analysis_list$char_mapped <- get_feature_summary(post_covid, 
                                                 features = c(ahlqvist, 
                                                              # "race", # workspace too small to calculate p for fisher.test
                                                              "gender",
                                                              # demo, # no deceased in SAID, causes error
                                                              complications,
                                                              glucose_homeostasis,
                                                              liver,
                                                              lipid,
                                                              kidney,
                                                              prescription),
                                                 categories = "mapped_cluster", 
                                                 distribution_stat = "ci") %>% feature_rename() %>% rename(`Unadjusted p-value` = `Mapped_cluster p-value`)
```

Distribution of cluster assignment appears different by gender, particularly in the SIDD cluster.

```{r}
post_covid %>% 
  calc_cat_summary(df = ., features = "mapped_cluster", categories = "gender") %>% 
  reorder_columns() %>% 
  arrange(factor(Value, levels = c("SAID", "SIDD", "SIRD", "MOD", "MARD")))
```

Distribution of cluster assignment does not appear different by race.

```{r}
post_covid %>% 
  filter(race != "Other") %>% 
  calc_cat_summary(df = ., features = "mapped_cluster", categories = "race") %>% 
  reorder_columns() %>% 
  arrange(factor(Value, levels = c("SAID", "SIDD", "SIRD", "MORD", "MARD")))
```

Cluster distribution is not even in White patients.

```{r}
chisq.test(table(white$Ahlqvist))
```

Cluster distribution is not even in Black/African American patients.

```{r}
chisq.test(table(aa$Ahlqvist))
```


# MARD vs other clusters
MARD appears to generally have the mildest phenotype.

```{r}
analysis_list$char_mard_v_others <- get_feature_summary(post_covid, 
                                                      features = c(ahlqvist, 
                                                                   "race", "gender", demo, 
                                                                   complications, 
                                                                   glucose_homeostasis, 
                                                                   liver, 
                                                                   lipid, 
                                                                   kidney, 
                                                                   prescription),
                                                      categories = c("MARD"),
                                                      covariates = covariates) %>%
  rename("Other" = "0",
         "MARD" = "1") %>% 
  feature_rename() %>% rename(`Unadjusted p-value` = `Mard p-value`, `Adjusted p-value` = `Mard adjusted p-value`)
```

## Demogrpahics

Logistic regression was used to determine if the odds of being assigned to the MARD cluster is significantly different with race. Covariates for regression model were chosen based on background knowledge.

```{r}
analysis_list$multi_mard_v_other_cluster <- logit_table(post_covid, 
                                                 "MARD", 
                                                 covariates) %>%
  rename("Other" = "0",
         "MARD" = "1") %>%
  arrange(factor(Value, levels = c("White", "Black or African American", "Other",
                                   "Male", "Female")))

```

## Complications
Aggregate complication risks.

```{r}
analysis_list$multi_mard_agg_comp <- logit_table(post_covid,
                                                 "complications_all",
                                                 c("MARD", covariates)) %>%
    rename("No" = "0",
           "Yes" = "1") %>%
    arrange(factor(Value, levels = c("Yes", "No",
                                     "White", "Black or African American", "Other",
                                     "Male", "Female"))) 

analysis_list$multi_mard_agg_comp_wo_nafld <- logit_table(post_covid,
                                                 "complications_except_nafld",
                                                 c("MARD", covariates)) %>%
    rename("No" = "0",
           "Yes" = "1") %>%
    arrange(factor(Value, levels = c("Yes", "No",
                                     "White", "Black or African American", "Other",
                                     "Male", "Female"))) 
```

Individual complication risks.

```{r}
odds <- c("u_ketone_first_positive", "mi", "neuropathy", "nephropathy", "retinopathy", "nafld")

for (o in odds){
  analysis_list[[paste("multi_mard", o, sep = "_")]] <- logit_table(post_covid, 
                                                                    o,
                                                                    c("MARD", covariates)) %>%
    rename("No" = "0",
           "Yes" = "1") %>%
    arrange(factor(Value, levels = c("Yes", "No",
                                     "White", "Black or African American", "Other",
                                     "Male", "Female"))) 
}

```

# AA vs White in MARD cluster

```{r}
mard_w_aa <- post_covid %>% filter((Ahlqvist == "MARD") & (race == "White" | race == "Black or African American"))

analysis_list$char_mard_w_v_aa <- get_feature_summary(mard_w_aa, 
                                                    features = c(ahlqvist[ahlqvist != "gad"], 
                                                                 "gender", demo, 
                                                                 complications, 
                                                                 glucose_homeostasis, 
                                                                 liver, 
                                                                 lipid, 
                                                                 kidney, 
                                                                 prescription),
                                                    categories = "race",
                                                    covariates = covariates[covariates != "race"]) %>% 
  feature_rename() %>% 
  rename(`Unadjusted p-value` = `Race p-value`, `Adjusted p-value` = `Race adjusted p-value`)

odds <- c("u_ketone_first_positive", "mi", "neuropathy", "nephropathy", "retinopathy", "nafld")

for (o in odds){

  analysis_list[[paste("multi_mard_waa", o, sep = "_")]] <- logit_table(mard_w_aa,
                                                                    o,
                                                                    covariates) %>%
    rename("No" = "0",
           "Yes" = "1") %>%
    arrange(factor(Value, levels = c("Yes", "No", 
                                     "White", "Black or African American", "Other",
                                     "Male", "Female")))

}
```

# SAID vs other clusters

Since SAID should be analogous to type 1 diabetes, insulin prescription should be higher in SAID compared to other clusters.

```{r}
# unadjusted
analysis_list$uni_or_said_insulin_prescription <- logit_table(post_covid, 
                                                                 "insulin_prescription", 
                                                                 "SAID") %>%
  rename("No insulin prescription" = "0",
         "Insulin prescription" = "1")
```

Similarly, urine ketone should also be higher. However, urine ketone was not significantly higher in SAID compared to other clusters.

```{r}
# unadjusted
analysis_list$uni_or_said_u_ketone <- logit_table(post_covid, 
                                                 "u_ketone_first_positive", 
                                                 "SAID") %>%
  rename("No urine ketone" = "0",
         "Urine ketone" = "1")

```

Since SIDD is defined by its relatively low beta cell function, urine ketone could also be higher in SIDD compared to other clusters, which could be masking the expected higher urine ketone in SAID. Indeed, the odds ratio of having a positive urine ketone in SAID was almost significant compared to all other clutsers except SIDD.

```{r}
no_sidd <- post_covid %>% filter(Ahlqvist != "SIDD")

analysis_list$uni_or_said_no_sidd_u_ketone <- logit_table(no_sidd, 
                                                 "u_ketone_first_positive", 
                                                 "SAID") %>%
  rename("No urine ketone" = "0",
         "Urine ketone" = "1")

```

SIDD and SAID have similar odds ratio of having a positive urine ketone test.

```{r}
said_sidd <- post_covid %>% filter(Ahlqvist == "SAID" | Ahlqvist == "SIDD")

analysis_list$uni_or_sidd_said_u_ketone <- logit_table(said_sidd, 
                                                 "u_ketone_first_positive", 
                                                 "SIDD") %>%
  rename("No urine ketone" = "0",
         "Urine ketone" = "1")
```

# SIDD vs other clusters

```{r}

analysis_list$char_sidd_v_others <- get_feature_summary(post_covid, 
                                                      features = c(ahlqvist, 
                                                                   "race", "gender", demo, 
                                                                   complications, 
                                                                   glucose_homeostasis, 
                                                                   liver, 
                                                                   lipid, 
                                                                   kidney, 
                                                                   prescription),
                                                      categories = c("SIDD"),
                                                      covariates = covariates) %>%
  rename("Other" = "0",
         "SIDD" = "1") %>% 
  feature_rename() %>% rename(`Unadjusted p-value` = `Sidd p-value`, `Adjusted p-value` = `Sidd adjusted p-value`)
```

## Demogrpahics

Logistic regression was used to determine if the odds of being assigned to the SIDD cluster is significantly different with race. Covariates for regression model were chosen based on background knowledge.

```{r}
analysis_list$multi_sidd_v_other_cluster <- logit_table(post_covid, 
                                                 "SIDD", 
                                                 covariates) %>%
  rename("Other" = "0",
         "SIDD" = "1") %>%
  arrange(factor(Value, levels = c("White", "Black or African American", "Other",
                                   "Male", "Female")))

analysis_list$multi_sidd_aa_factor <- logit_table(post_covid, 
                                                 "aa_factor", 
                                                 c("SIDD", "gender", "diabetes_age")) %>%
  rename("Other" = "0",
         "AA" = "1")

```

## Prescription

While the high percentage of patients with insulin prescription in SAID is expected, the percentage of patients with insulin prescription was also higher in SIDD compared to other clusters.

```{r}
analysis_list$multi_sidd_insulin_prescription <- logit_table(post_covid, 
                                                                 "insulin_prescription", 
                                                                 c("SIDD", covariates)) %>%
  rename("No insulin prescription" = "0",
         "Insulin prescription" = "1")

analysis_list$multi_sidd_said_insulin_prescription <- logit_table(said_sidd, 
                                                                 "insulin_prescription", 
                                                                 c("SIDD", covariates)) %>%
  rename("No insulin prescription" = "0",
         "Insulin prescription" = "1")
```

## SIDD complications
Aggregate complication risks.

```{r}
analysis_list$multi_sidd_agg_comp <- logit_table(post_covid,
                                                 "complications_all",
                                                 c("SIDD", covariates)) %>%
    rename("No" = "0",
           "Yes" = "1") %>%
    arrange(factor(Value, levels = c("Yes", "No",
                                     "White", "Black or African American", "Other",
                                     "Male", "Female"))) 

analysis_list$multi_sidd_agg_comp_wo_nafld <- logit_table(post_covid,
                                                 "complications_except_nafld",
                                                 c("SIDD", covariates)) %>%
    rename("No" = "0",
           "Yes" = "1") %>%
    arrange(factor(Value, levels = c("Yes", "No",
                                     "White", "Black or African American", "Other",
                                     "Male", "Female"))) 
```

Individual complication risks.

While the odds of patients in the SIDD cluster with a positive urine ketone test was higher than that of other clusters, the odds of being diagnosed with DKA was not significantly higher.

SIDD patients have higher odds of neuropathy, MI, and urine ketone. 

```{r}
for (comp in complications){
  # unadjusted
  analysis_list[[paste("uni_or_sidd", comp, sep = "_")]] <- logit_table(post_covid, 
                                                                        comp, 
                                                                        "SIDD") %>%
    rename("No" = "0",
           "Yes" = "1")
  
  # adjusted for demographic covariates
  analysis_list[[paste("multi_sidd", comp, sep = "_")]] <- logit_table(post_covid, 
                                                                       comp, 
                                                                       c("SIDD", covariates)) %>%
    rename("No" = "0",
           "Yes" = "1") %>%
    arrange(factor(Value, levels = c("Yes", "No", 
                                     "White", "Black or African American", "Other",
                                     "Male", "Female")))
  
   }
```

# White vs AA in SIDD cluster

```{r}
w_aa <- post_covid %>% filter((Ahlqvist == "SIDD") & (race == "White" | race == "Black or African American"))

analysis_list$char_sidd_w_v_aa <- get_feature_summary(w_aa, 
                                                    features = c(ahlqvist[ahlqvist != "gad"], 
                                                                 "gender", demo, 
                                                                 complications, 
                                                                 glucose_homeostasis, 
                                                                 liver, 
                                                                 lipid, 
                                                                 kidney, 
                                                                 prescription),
                                                    categories = "race",
                                                    covariates = covariates[covariates != "race"]) %>% 
  feature_rename() %>% 
  rename(`Unadjusted p-value` = `Race p-value`, `Adjusted p-value` = `Race adjusted p-value`)
```

## Race and Complications
Logistic regression to determine odds ratios.

```{r}
odds <- c("u_ketone_first_positive", "mi", "neuropathy", "nephropathy", "retinopathy", "nafld")

for (o in odds){

  analysis_list[[paste("multi_sidd_waa", o, sep = "_")]] <- logit_table(w_aa,
                                                                    o,
                                                                    covariates) %>%
    rename("No" = "0",
           "Yes" = "1") %>%
    arrange(factor(Value, levels = c("Yes", "No", 
                                     "White", "Black or African American", "Other",
                                     "Male", "Female")))

}

```

# Calculate n for figures

```{r}
analysis_list$figure1_n <- post_covid %>% 
  group_by(Ahlqvist) %>%
  summarize(n = n())

analysis_list$figure2_n <- post_covid %>% 
  group_by(Ahlqvist) %>%
  summarize(male = sum(gender == "Male"),
            female = sum(gender == "Female"),
            white = sum(race == "White"),
            AA = sum(race == "Black or African American"))

analysis_list$figure3_n <- post_covid %>% 
  group_by(SIDD) %>%
  summarize(homa2b = sum(!is.na(homa_cpeptide_b)),
            cpeptide = sum(!is.na(cpeptide)),
            glucose = sum(!is.na(glucose)),
            a1c = sum(!is.na(a1c)),
            non_hdl = sum(!is.na(non_hdl)),
            triglyceride_hdl = sum(!is.na(triglyceride_hdl)),
            diabetes_age = sum(!is.na(diabetes_age)),
            deceased_age = sum(!is.na(deceased_age)))

analysis_list$figure4_n <- w_aa %>% 
  group_by(race) %>%
  summarize(homa2b = sum(!is.na(homa_cpeptide_b)),
            cpeptide = sum(!is.na(cpeptide)),
            glucose = sum(!is.na(glucose)),
            a1c = sum(!is.na(a1c)),
            bmi = sum(!is.na(bmi)),
            homa2ir = sum(!is.na(homa_cpeptide_ir)),
            non_hdl = sum(!is.na(non_hdl)),
            triglyceride_hdl = sum(!is.na(triglyceride_hdl)))

analysis_list$figure5_n <- post_covid %>% 
  group_by(MARD) %>%
  summarize(diabetes_age = sum(!is.na(diabetes_age)),
            deceased_age = sum(!is.na(deceased_age)),
            glucose = sum(!is.na(glucose)),
            a1c = sum(!is.na(a1c)),
            bmi = sum(!is.na(bmi)),
            homa2ir = sum(!is.na(homa_cpeptide_ir)),
            non_hdl = sum(!is.na(non_hdl)),
            triglyceride_hdl = sum(!is.na(triglyceride_hdl)))

analysis_list$text_homa2_n <- homa_ic %>% group_by(method) %>% summarize(homa2b = length(homa2b),
                                                                   homa2ir = length(homa2ir),
                                                                   White = sum(race == "White"),
                                                                   AA = sum(race == "Black or African American"),
                                                                   Other = sum(race == "Other"),
                                                                   Male = sum(gender == "Male"),
                                                                   Female = sum(gender == "Female"))
  
```


# Export results
Create characteristic table for Supplemental Table S1.

```{r}

s1_feature_order <- c("Gender", "Race",
                      "HbA1c", "BMI", "Diagnosis age", 
                      "HOMA2-B", "HOMA2-IR", "Glucose", "C-peptide", 
                      "Cholesterol", "LDL", "HDL", "Triglycerides", 
                      "Urine albumin")

s1_features <- c("gender", "race",
                 "bmi", "diabetes_age", 
                 "homa_cpeptide_b", "glucose", 
                 "cholesterol", "ldl", "hdl", "triglyceride", 
                 "u_albumin")

s1_features_decimal <- c("a1c", "homa_cpeptide_ir", "cpeptide")


table_s1_int <- get_feature_summary(post_covid, s1_features, "Ahlqvist", distribution_stat = "ci", digits = 0) %>% 
  feature_rename() %>% 
  rename(P = `Ahlqvist p-value`)
table_s1_decimal <- get_feature_summary(post_covid, s1_features_decimal, "Ahlqvist", distribution_stat = "ci", digits = 1) %>% 
  feature_rename() %>% 
  rename(P = `Ahlqvist p-value`)

table_s1 <- full_join(table_s1_int, table_s1_decimal)

table_s1 <- table_s1 %>% 
  select(Features, Value, SAID, SIDD, SIRD, MOD, MARD, P) %>% # subset to relevant columns 
  filter(!grepl("No|Female|White|Other", .$Value))

# rearrange rows to order supplied
table_s1$Features <- factor(table_s1$Features, levels = s1_feature_order)
table_s1 <- table_s1 %>% arrange(Features)

# format Feature with units
table_s1 <- table_s1 %>%
  mutate(Features = case_when(Features == "Gender" ~ "Male",
                              Features == "Race" ~ "AA",
                              .default = Features),
         Value = case_when(Value == "Yes" ~ "(% pos)", # change unit values
                           Value == "Years" ~ "(years)",
                           Value == "kg/m^2" ~ "(kg/m2)",
                           Features == "HOMA2-B" | Features == "HOMA2-IR" ~ "",
                           Features == "Male" | Features == "AA" ~ "(%)",
                           .default = paste0("(", Value, ")"))) %>%
  mutate(Features = trimws(paste(Features, Value))) %>% # paste units to feature names
  select(-Value) # remove unit column

# add line break
for (c in order){
  #table_s1[[c]] <- str_replace_all(table_s1[[c]], "\\d* \\((\\d*)\\)", "\\1")
  table_s1[[c]] <- str_replace_all(table_s1[[c]], ", ", "-")
  table_s1[[c]] <- str_replace_all(table_s1[[c]], " \\(", "\n\\(")
}

# add counts and frequency row
s1_n <- post_covid %>% count(Ahlqvist) %>% pivot_wider(names_from = Ahlqvist, values_from  = n)
s1_freq <- as_tibble(round(prop.table(s1_n) * 100, 0)) %>% mutate(across(everything(), as.character)) %>% mutate(Features = "Frequency (%)")
s1_n <- s1_n %>% mutate(across(everything(), as.character)) %>% mutate(Features = "N")
table_s1 <- bind_rows(s1_n, s1_freq, table_s1)

# reorder columns
table_s1 <- table_s1 %>% select(Features, {{order}}, P) %>% 
  rename(Clusters = Features) %>%
  mutate(P = case_when(as.numeric(P) > 0.05 ~ paste0("(", as.numeric(P), ")"), .default = P)) # parenthesis p > 0.05

```

Combine univariate analysis tables into Excel workbook.

```{r}
# create workbook
wb <- createWorkbook()

# define sheet styles
header_style <- createStyle(textDecoration = "Bold", border = c("top", "bottom", "left", "right") , halign = "center", valign = "center")
sheet_style <- createStyle(border = c("top", "bottom", "left", "right"), halign = "center", valign = "center")

# define names to loop through
analysis_names <- c("char_ahlqvist", "char_mard_v_others", "char_sidd_v_others", "char_sidd_w_v_aa")
excel_names <- c("Cluster Characteristics", "MARD vs Others", "SIDD vs Others", "SIDD AA vs White")


for (i in 1:length(analysis_names)){
  # rearrange race rows
  if ("Race" %in% analysis_list[[analysis_names[i]]]$Features){
    # find rows
    white_index <- match("White", analysis_list[[analysis_names[i]]]$Value)
    aa_index <- match("Black or African American", analysis_list[[analysis_names[i]]]$Value)
    
    # slice out row with AA
    aa_row <- slice(analysis_list[[analysis_names[i]]], aa_index)
    
    # remove row from original and add back in correct position
    analysis_list[[analysis_names[i]]] <- analysis_list[[analysis_names[i]]] %>% 
      filter(row_number() != aa_index) %>% 
      add_row(aa_row, .after = white_index)
  }
  
  # remove duplicated features to merge
  dup <- duplicated(analysis_list[[analysis_names[i]]]$Features)
  excel_formatted <- analysis_list[[analysis_names[i]]] %>% mutate(Features = case_when(dup ~ NA, .default = Features),
                    Summary = case_when(dup ~ NA, .default = Summary),
                    `Unadjusted p-value` = case_when(dup ~ NA, .default = `Unadjusted p-value`))
  
  # create and write excel sheet
  addWorksheet(wb, sheetName = excel_names[i], gridLines = F)
  writeData(wb, sheet = excel_names[i], x = excel_formatted, borders = "all", headerStyle = header_style)
  
  # find features that span multiple rows
  dup_features <- analysis_list[[analysis_names[i]]] %>% count(across(Features)) %>% filter(n > 1) %>% pull(Features)
  
  # find indices of columns to merge
  merge_cols <- which(colnames(analysis_list[[analysis_names[i]]]) %in% c("Features", "Summary", "Unadjusted p-value"))
  
  # loop to merge
  for (f in dup_features){
    # find starting and ending row indices containing feature
    dup_index <- which(analysis_list[[analysis_names[i]]]$Features == f)
    first_index <- min(dup_index)+1
    last_index <- max(dup_index)+1
    
    for (c in merge_cols){
      # merge columns containing feature name, summary, and p-value
      mergeCells(wb, excel_names[i], cols = c, rows = first_index:last_index)
      
    }
    
  }
  
  # center align entire sheet
  addStyle(wb, excel_names[i], sheet_style, 
           rows = 1:nrow(excel_formatted) + 1, # start at 1 to skip header row
           cols = 1:ncol(excel_formatted), gridExpand = T)
  
  # set column width
  setColWidths(wb, excel_names[i], cols = 1:ncol(excel_formatted), widths = "auto")
}


# add supplemental table 1
addWorksheet(wb, sheetName = "Table S1", gridLines = F)
writeData(wb, sheet = "Table S1", x = table_s1, borders = "none", 
          headerStyle = createStyle(textDecoration = "Bold", border = c("top", "bottom"), halign = "center", valign = "center"))
# bold first column
addStyle(wb, "Table S1", createStyle(textDecoration = "Bold", halign = "left", valign = "center"), 
         rows = 2:(nrow(table_s1) + 1), # skip header row
         cols = 1, gridExpand = T)
# fix formatting for first cell
# addStyle(wb, "Table S1", createStyle(textDecoration = "Bold", halign = "left", valign = "center", border = c("top", "bottom")), 
#          rows = 1, 
#          cols = 1, gridExpand = T)
# center align sheet except first and last column
addStyle(wb, "Table S1", createStyle(textDecoration = NULL, halign = "center", valign = "center"), 
         rows = 1:nrow(table_s1) + 1, # skip header row
         cols = 2:ncol(table_s1), gridExpand = T) # except first column
# italicize p-value
addStyle(wb, "Table S1", createStyle(textDecoration = "italic", halign = "center", valign = "center", fontSize = 9), 
         rows = 3:nrow(table_s1) + 1, # skip header row
         cols = ncol(table_s1), gridExpand = T)
# set column width
setColWidths(wb, "Table S1", cols = 1:ncol(table_s1), widths = "auto")
```


