---
title: "analysis_nhanes"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Environment setup

Import libraries

```{r}
library(tidyverse)
library(survey)
library(openxlsx)
```

Set root directory

```{r}
here::i_am("README.md")
```

## Import data
Cluster data is imported from

```{r}

export_dir <- here::here("exports", "NHANES")
nhanes_clusters <- read_csv(file.path(export_dir, "nhanes_denovo_clusters.csv"), show_col_types = F)
nhanes_data <- read_rds(file.path(export_dir, "nhanes_raw_data.rds"))

```


# Data cleaning

Join NHANES clustering data with NHANES data.

```{r}
data <- nhanes_clusters %>% 
  full_join(nhanes_data, by = join_by(SEQN))
```

Use WTSAF2YR and WTSAFPRP for weight calculations since fasting blood samples used. Cycle P is 3.2 years, rest is 2 years

```{r}

nhanes_cycles <- unique(data$cycle)
cycle2yr <- nhanes_cycles[nhanes_cycles != "P"]
sum_years <- length(cycle2yr)*2 + length(nhanes_cycles[nhanes_cycles == "P"])*3.2 # Cycle P is 3.2 years, rest is 2 years

data <- data %>%
  mutate(adjusted_mec_weights = case_when(cycle %in% cycle2yr ~ WTSAF2YR * (2/sum_years),
                                          cycle == "P" ~ WTSAFPRP * (3.2/sum_years)))
```

Set factor levels for analysis.

```{r}

data <- data %>%
  mutate(RIDRETH1 = relevel(factor(RIDRETH1), ref = "Non-Hispanic White"),
         race = relevel(factor(case_when(RIDRETH1 == "Non-Hispanic White" ~ "White", 
                          RIDRETH1 == "Non-Hispanic Black" ~ "AA",
                          .default = "Other")), ref = "White"),
         overall = factor("Overall"),
         # adjusted_diet_weights = case_when(is.na(adjusted_diet_weights) ~ 0, .default = adjusted_diet_weights),
         adjusted_mec_weights = case_when(is.na(adjusted_mec_weights) ~ 0, .default = adjusted_mec_weights),
         sidd = case_when(hier_clusters == "SIDD" ~ 1, is.na(hier_clusters) ~ NA, .default = 0),
         sird = case_when(hier_clusters == "SIRD" ~ 1, is.na(hier_clusters) ~ NA, .default = 0),
         mod = case_when(hier_clusters == "MOD" ~ 1, is.na(hier_clusters) ~ NA, .default = 0),
         mard = case_when(hier_clusters == "MARD" ~ 1, is.na(hier_clusters) ~ NA, .default = 0),
         SIDD = factor(case_when(hier_clusters == "SIDD" ~ "SIDD", is.na(hier_clusters) ~ NA, .default = "Others"), levels = c("Others", "SIDD")),
         severe = case_when(hier_clusters == "SIDD" | hier_clusters == "SIRD" ~ 1, 
                            hier_clusters == "MOD" | hier_clusters == "MARD" ~ 0),
         Severe = case_when(severe == 1 ~ "Severe", severe == 0 ~ "Mild"),
         pandemic = factor(case_when(cycle == "L" ~ "2021-2023", cycle == "I" | cycle == "P" ~ "2015-2020")),
         # insulin use
         diagnosed_diabetes = case_when(DIQ010 == "Yes" ~ 1, .default = 0),
         diabetes = case_when(DIQ010 == "Yes" ~ 1,
                              LBXGH >= 6.5 ~ 1,
                              LBDGLUSI >= 7 ~ 1,
                              .default = 0),
         insulin_med = case_when(DIQ050 == "Yes" ~ 1, 
                                 DIQ050 == "No" ~ 0),
         insulin_med_cat = case_when(insulin_med == 1 ~ "Yes", 
                                     insulin_med == 0 ~ "No"),
         diabetes_pill = case_when(DIQ070 == "Yes" ~ 1,
                                   DIQ070 == "No" ~ 0),
         diabetes_pill_cat = case_when(diabetes_pill == 1 ~ "Yes",
                                       diabetes_pill == 0 ~ "No"),
         insurance = case_when(HIQ011 == "Yes" ~ 1, 
                               HIQ011 == "No" ~ 0))
         

```

Set overall design. N in design doesn't match all available respondents with clusters since some respondents have an adjusted weight of 0.

```{r}
NHANES_all <- svydesign(data=data, id=~SDMVPSU, strata=~SDMVSTRA, weights=~adjusted_mec_weights, nest=TRUE)

NHANES <- subset(NHANES_all, 
                  diabetes == 1 &
                   RIDAGEYR >= 18 & 
                   (DID040 >= 18 | is.na(DID040)) & 
                   !is.na(hier_clusters))

```

# Analysis

Source analysis functions.

```{r}
# import functions to summarize features into tables
source("/data/user/brianlu/projects/precision_diabetes/Clustering/functions_analysis.R")

# import functions to analyze survey data
source(here::here("NHANES", "survey_analysis_functions.R"))
```


```{r}

# subset to participants used and with non-zero weight
data_in_analysis <- data %>% filter(diabetes == 1 &
                                      RIDAGEYR >= 18 & 
                                      (DID040 >= 18 | is.na(DID040)) & 
                                      !is.na(hier_clusters)& 
                                      adjusted_mec_weights > 0)

# construct patient_char tibble
patient_n = table(data_in_analysis$pandemic)
patient_char <- tibble(Period = c("Sex, No. (%)", "Race, No. (%)", "Diabetes subtype, No. (%)", "Severe subtype, No. (%)"), 
                       pre = "",
                       post = "",
                       P = as.numeric(NA))


#######################
# construct counts table to join with patient_char

# calculate counts for table
baseline <- get_feature_summary(df = data_in_analysis, 
                                features = c("RIAGENDR", "race", "hier_clusters", "Severe"),
                                categories = c("pandemic"),
                                covariates = NA,
                                distribution_stat = "ci")

# calculate weighted percentages
cat_summary <- map(c(~RIAGENDR, ~race, ~hier_clusters, ~Severe), 
                   get_cat_summary, 
                   byformula = ~pandemic, 
                   design = NHANES) %>% 
  bind_rows()

# calculate p-values
cat_p <- c(rep(svychisq(~RIAGENDR + pandemic, NHANES, statistic = "adjWald")$p.value, 2), # twice for male/female
           rep(svychisq(~race + pandemic, NHANES, statistic = "adjWald")$p.value, 3), # 3 for White, AA, and Other
           rep(svychisq(~hier_clusters + pandemic, NHANES, statistic = "adjWald")$p.value, 4), # 4 for 4 clusters
           rep(svychisq(~Severe + pandemic, NHANES, statistic = "adjWald")$p.value, 2)) # 2 for mild/severe


counts_table <- baseline %>% 
  inner_join(cat_summary, by = join_by(Features, Value)) %>%
  mutate(pre = str_replace(`2015-2020.x`, "\\(\\d+\\)", paste0("(", `2015-2020.y`, ")")),
         post = str_replace(`2021-2023.x`, "\\(\\d+\\)", paste0("(", `2021-2023.y`, ")"))) %>%
  # filter(Value %in% c("Male", "AA", "White")) %>%
  mutate(Period = paste0("  ", Value)) %>% # rename columns to match and indent
  mutate(P = cat_p) %>% # add svychisq p-values
  select(Period, pre, post, P) # select joining columns

# constructs continuous table to join with patient_char

cont_variables <- c(#"LBXGH", # a1c %
                    "BMXBMI", #BMI kg/m2
                    "RIDAGEYR")#, # age year
                    # "homa_insulin_ir", 
                    # "LBDGLUSI", # glucose mM
                    # "LBDINSI",  #insulin pM
                    # "LBXTC", # total cholesterol mg/dL
                    # "LBDHDD") # HDL mg/dL

# calculate p-values from svyttest
cont_p <- purrr::map(purrr::map2("pandemic", cont_variables, reformulate),
                         svyttest, design = NHANES) %>%
  purrr::map_vec(., (function(x) x$p.value))

# build table
cont_summary <- map(map(cont_variables, reformulate), 
                   get_cont_summary, 
                   byformula = ~pandemic, 
                   design = NHANES) %>% 
  bind_rows() %>% 
  mutate(P = cont_p) %>%
  mutate(Features = case_when(Features == "LBXGH" ~ "HbA1c [%]",
                              Features == "BMXBMI" ~ "BMI, mean (95% CI), kg/m2",
                              Features == "RIDAGEYR" ~ "Age, mean (95% CI), y",
                              Features == "homa_insulin_ir" ~ "HOMA2-IR",
                              Features == "LBDGLUSI" ~ "Glucose [mmol/L]",
                              Features == "LBDINSI" ~ "Insulin [pmol/L",
                              Features == "LBXTC" ~ "Cholesterol [mg/dL]",
                              Features == "LBDHDD" ~ "HDL [mg/dL]")
         ) %>%
  rename(Period = Features,
         pre = `2015-2020`,
         post = `2021-2023`)


clusters <- c("SIDD", "SIRD", "MOD", "MARD")

table2 <- bind_rows(patient_char, counts_table, cont_summary) %>%
  mutate(P = sprintf("%.3f", P),
         P = str_replace(P, "NA", "")) %>%
  rename("2015-2020\n(n = {patient_n[1]})" := pre,
         "2021-2023\n(n = {patient_n[2]})" := post) %>%
  
   mutate(Period = factor(Period, levels = c("Sex, No. (%)", 
                                             "  Male", "  Female",
                                             "Race, No. (%)", "  White", "  AA", "  Other", 
                                             "Age, mean (95% CI), y", 
                                             "BMI, mean (95% CI), kg/m2",
                                             "Diabetes subtype, No. (%)", paste0("  ", clusters),
                                             "Severe subtype, No. (%)", "  Severe", "  Mild"))) %>%
  arrange(Period) #rearrange order by factor

```

Add formatting and export as excel file.

```{r}
# create workbook
wb <- createWorkbook()

sheet_name <- "NHANES Char Table"

# add table 1
addWorksheet(wb, sheetName = sheet_name, gridLines = F)
writeData(wb, sheet = sheet_name, x = table2, borders = "all", 
          headerStyle = createStyle(textDecoration = "Bold", 
                                    border = c("top", "bottom", "left", "right"), 
                                    halign = "left", 
                                    valign = "center",
                                    wrapText = TRUE))

# set column width
setColWidths(wb, sheet_name, cols = 1:ncol(table2), widths = "auto")

# export
saveWorkbook(wb, file.path(export_dir, "nhanes_denovo_char_table.xlsx"), overwrite = TRUE)
```

## Cluster features

```{r}

# define plot order and colors
order <- c("SIDD", "SIRD", "MOD", "MARD")
color <- c("powderblue", "darkseagreen", "lightsalmon", "plum")

# define plotting features to loop
y_data <- c("LBXGH", 
            "RIDAGEYR",
            "BMXBMI", 
            "homa1_cpeptide_b",
            "homa1_cpeptide_ir")

y_labels <- list("HbA1c (%)",
                 "Age (years)",
                 bquote("BMI " (kg/m^2)),
                 expression(atop("Scaled", paste("HOMA1"["insulin"], "-B", " (%)"))),
                 expression(atop("Scaled", paste("HOMA1"["insulin"], "-IR"))))

names(y_labels) <- y_data

# import clustering features
clustering_features <- read_csv(file.path(export_dir, "nhanes_clustering_features.csv"), show_col_types = F)

# add cluster labels
clustering_features <- clustering_features %>%
  inner_join(nhanes_clusters, by = join_by(SEQN)) %>%
  filter(SEQN %in% data_in_analysis$SEQN)

# graph
for (y in y_data){
 feature_plot <- ggplot(clustering_features, mapping = aes(x = hier_clusters, y = .data[[y]], fill = hier_clusters)) + 
                        stat_boxplot(geom = "errorbar", width = 0.25, linewidth = 0.25) +
                        geom_boxplot(outlier.size = 0.25, show.legend = F, linewidth = 0.25) + 
                        scale_fill_manual(values = color, limits = order) +
                        scale_x_discrete(limits = order) + 
                        theme_classic() + 
                        labs(y = y_labels[[y]] , x = "") + 
                        theme(axis.text.x = element_text(angle = 45, vjust = 0.5),
                              plot.margin = margin(t = 13)) 
  
  ggsave(file.path(export_dir, paste0("nhanes_denovo_feature_", y, ".tiff")), feature_plot, dpi = 600, 
         width = ifelse(str_detect(y, "homa"), 4.0, 3.5), 
         height = 4.5, units = "cm")
}
```


## Severe subtype regression

```{r}

covariates <- c("RIAGENDR", "race", #"RIDRETH1", 
                "RIDAGEYR", "BMXBMI", "diagnosed_diabetes", "pandemic")

severe_logit <- get_survey_or("severe", covariates, NHANES)

write_csv(severe_logit, file.path(export_dir, "nhanes_denovo_severe_regression.csv"))

```
## Subtype vs others regression

```{r}

subtypes <- c("sidd", "sird", "mod", "mard")

for (s in subtypes){
  s_vs_others <- get_survey_or(s, covariates, NHANES)
  
  write_csv(s_vs_others, file.path(export_dir, paste("nhanes_denovo", s, "vs_rest_regression.csv", sep = "_")))
}

```

## Insulin prescription

Construct insulin table.

```{r}
# subset data to participants with valid insulin question answers
NHANES_insulin <- subset(NHANES, !is.na(insulin_med_cat))
data_insulin_analysis <- data_in_analysis %>% filter(!is.na(insulin_med_cat))

insulin_char <- tibble(Subtype = c("Taking insulin, No. (%)"), 
                       Overall = "",
                       SIDD = "",
                       SIRD = "",
                       MOD = "",
                       MARD = "",
                       P = as.numeric(NA))

baseline <- get_feature_summary(df = data_insulin_analysis, 
                                features = "insulin_med_cat",
                                categories = c("hier_clusters"),
                                covariates = NA,
                                distribution_stat = "ci")

# calculate weighted percentages
cat_summary <- map(c(~insulin_med_cat), 
                   get_cat_summary, 
                   byformula = ~hier_clusters, 
                   design = NHANES_insulin) %>% 
  bind_rows()

overall_summary <- get_cat_summary(~insulin_med_cat, ~overall, NHANES_insulin)

cat_summary <- cat_summary %>%
  full_join(overall_summary, by = join_by(Features, Value))

# calculate p-values
cat_p <- rep(svychisq(~hier_clusters + insulin_med_cat, NHANES_insulin, statistic = "adjWald")$p.value, 2) # 2 for yes and no


counts_table <- baseline %>% 
  inner_join(cat_summary, by = join_by(Features, Value)) %>%
  mutate(Overall = str_replace(Overall.x, "\\(\\d+\\)", paste0("(", Overall.y, ")")),
         SIDD = str_replace(SIDD.x, "\\(\\d+\\)", paste0("(", SIDD.y, ")")),
         SIRD = str_replace(SIRD.x, "\\(\\d+\\)", paste0("(", SIRD.y, ")")),
         MOD = str_replace(SIDD.x, "\\(\\d+\\)", paste0("(", MOD.y, ")")),
         MARD = str_replace(SIDD.x, "\\(\\d+\\)", paste0("(", MARD.y, ")"))) %>%
  # filter(Value %in% c("Male", "AA", "White")) %>%
  mutate(Subtype = paste0("  ", Value)) %>% # rename columns to match and indent
  mutate(P = cat_p) %>% # add svychisq p-values
  select(Subtype, Overall, SIDD, SIRD, MOD, MARD, P) # select joining columns


insulin_char_table <- bind_rows(insulin_char, counts_table) %>%
  mutate(P = case_when(P < 0.001 ~ "<0.001", 
                       .default = sprintf("%.3f", P)),
         P = str_replace(P, "NA", ""))

write_csv(insulin_char_table, file.path(export_dir, "nhanes_denovo_insulin_char.csv"))
```

Also check if insulin use increased after pandemic.

```{r}

insulin_logit <- get_survey_or("insulin_med", c("hier_clusters", covariates), NHANES_insulin)

write_csv(insulin_logit, file.path(export_dir, "nhanes_denovo_insulin_regression.csv"))

```

## Cluster comparision to full *de novo* clustering

Mapped cluster assignments have moderate agreement to *de novo* clustering assignments. Kappa interpretation is as suggested by [McHugh](https://doi.org/10.11613/BM.2012.031) instead of the more lenient [Landis and Koch](https://www.ncbi.nlm.nih.gov/books/NBK92287/table/executivesummary.t2/?report=objectonly).

```{r}
# read in mapped clustering data
nhanes_mapped_clusters <- read_csv(file.path(export_dir, "nhanes_mapped_clusters.csv"), show_col_types = F)

# keep only participants with both assignments
cluster_combo <- inner_join(nhanes_clusters, nhanes_mapped_clusters, by = join_by(SEQN))


caret::confusionMatrix(reference = factor(cluster_combo$mapped_cluster), 
                       data = factor(cluster_combo$hier_clusters))
```

## Cluster comparison to *de novo* clustering separated by diabetes diagnosis status

*De novo* cluster assignment of entire NHANES data has moderate agreement with *de novo* cluster assignment separated by diabetes diagnosis status.

```{r}
diagnosed_cluster <- read_csv(file.path(export_dir, "nhanes_denovo_diagnosed_clusters.csv"), show_col_types = F) %>%
  mutate(diagnosed = "Yes")
undiagnosed_cluster <- read_csv(file.path(export_dir, "nhanes_denovo_undiagnosed_clusters.csv"), show_col_types = F) %>%
  mutate(diagnosed = "No")

diagnosis_cluster <- bind_rows(list(diagnosed_cluster, undiagnosed_cluster)) %>%
  rename(diagnosis_clusters = hier_clusters)
  
# keep only participants with both assignments
diagnosis_cluster_combo <- inner_join(nhanes_clusters, diagnosis_cluster, by = join_by(SEQN)) %>%
  inner_join(nhanes_mapped_clusters, by = join_by(SEQN))

caret::confusionMatrix(reference = factor(diagnosis_cluster_combo$hier_clusters), 
                       data = factor(diagnosis_cluster_combo$diagnosis_clusters))

```

*De novo* cluster assignment of medically diagnosed diabetes has moderate agreement with mapped cluster assignment.

```{r}

diagnosed <- diagnosis_cluster_combo %>% filter(diagnosed == "Yes")

caret::confusionMatrix(reference = factor(diagnosed$mapped_cluster), 
                       data = factor(diagnosed$diagnosis_clusters))

```

*De novo* cluster assignment of undiagnosed diabetes has substantial agreement with mapped cluster assignment.

```{r}

undiagnosed <- diagnosis_cluster_combo %>% filter(diagnosed == "No")

caret::confusionMatrix(reference = factor(undiagnosed$mapped_cluster), 
                       data = factor(undiagnosed$diagnosis_clusters))
```

Similarly, mapped cluster have higher agreement with undiagnosed diabetes than with diagnosed diabetes using clusters assigned using *de novo* clustering of the whole dataset.

```{r}

caret::confusionMatrix(reference = factor(diagnosed$mapped_cluster), 
                       data = factor(diagnosed$hier_clusters))
```

```{r}

caret::confusionMatrix(reference = factor(undiagnosed$mapped_cluster), 
                       data = factor(undiagnosed$hier_clusters))
```

# Session information

```{r}
sessionInfo()
```

