---
title: "Feature extraction for NHANES respondents with medically diagnosed diabetes"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NHANES clusters

Use [NHANES](https://wwwn.cdc.gov/nchs/nhanes/default.aspx) data to generate diabetes subtype clusters representative of the US population.

**Objective:** Extract required data from NHANES to generate diabetes clusters for participants.

# Environment setup

Load required libraries.

```{r}
library(tidyverse)
```

Set root directory

```{r}
here::i_am("README.md")
```

# Import data

Import previously downloaded NHANES data from the 2015-2016, 2017-2020 pre-pandemic, and 08/2021-08/2023 NHANES cycles.

```{r}

export_dir <- here::here("exports", "NHANES")
nhanes_data <- read_rds(here::here(export_dir, "nhanes_raw_data.rds"))

```

# Data cleaning

Filter data to only adults with medically diagnosed diabetes or adults meeting diagnostic criteria for diabetes. Respondents with insulin or glucose values outside the HOMA calculation range were excluded. Respondents with missing values were excluded.

```{r}
clustering_features <- nhanes_data %>%
  filter(RIDAGEYR >= 18) %>% # must be adult during survey
  filter((DIQ010 == "Yes" & DID040 >= 18 & DID040 <= 80) | DIQ010 != "Yes") %>% # must be adults at time of diagnosis if diagnosed
  # age above 80 coded as 80, values above 80 indicates unknown, less than 1 years old, or refused to answer
  filter(DIQ010 == "Yes" | LBXGH >= 6.5 | LBDGLUSI >= 7) %>% # must be diagnosed diabetes or meet diagnostic criteria
  filter(LBDINSI >= 20 & LBDINSI <= 400) %>% # filter insulin to HOMA range
  filter(LBDGLUSI >= 3.5 & LBDGLUSI <= 25) %>% # filter glucose to HOMA range
  select(SEQN, cycle, RIDRETH1, RIAGENDR, DIQ010, LBXGH, BMXBMI, RIDAGEYR, LBDINSI, LBDGLUSI) %>% # select cycle and required clustering features
  drop_na() # drop rows with missing values

```

Calculate insulinn HOMA estimates and convert to c-peptide equivalents.

```{r}

# import HOMA regression models
homa1b_model <- readRDS(here::here("exports/HOMA_regression/homa1b_model.RDS"))
homa1ir_model <- readRDS(here::here("exports/HOMA_regression/homa1ir_model.RDS"))

# calculate HOMA estimates
homa1 <- clustering_features %>%
  mutate(insulin_homa1b = (20 * LBDINSI/6) / (LBDGLUSI - 3.5),
         insulin_homa1ir = (LBDGLUSI * LBDINSI/6) / 22.5)

# rename columns to match regression model
homa1 <- homa1 %>%
  mutate(race = relevel(factor(case_when(RIDRETH1 == "Mexican American" ~ "Hispanic or Latino",
                          RIDRETH1 == "Other Hispanic" ~ "Hispanic or Latino",
                          .default = RIDRETH1)), ref = "Non-Hispanic White")) %>%
  rename(gender = RIAGENDR)

# re.form = NA to specify no random effects
homa1 <- homa1 %>%
  mutate(homa1_cpeptide_b = exp(predict(homa1b_model, newdata = homa1, re.form = NA)), 
         homa1_cpeptide_ir = exp(predict(homa1ir_model, newdata = homa1, re.form = NA)))

clustering_features <- clustering_features %>%
  mutate(homa1_cpeptide_b = homa1$homa1_cpeptide_b,
         homa1_cpeptide_ir = homa1$homa1_cpeptide_ir) %>% 
  select(-RIDRETH1, -RIAGENDR)
  
```

# Export data

Export the resulting data for clustering.

```{r}
# export for HOMA calculation
write_csv(clustering_features, file.path(export_dir, "nhanes_clustering_features.csv"))

```

# Session information

```{r}
sessionInfo()
```

