---
title: "Data extraction for NHANES"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Environment setup

Import libraries

```{r}
library(tidyverse)
library(nhanesA)
```

Set root directory

```{r}
here::i_am("README.md")
```

# Import data

NHANES data was imported using [nhanesA](https://doi.org/10.1093/database/baae028). Imported demographic and diabetes questionnaires, glycohemoglobin (HbA1c), insulin, and plasma fasting glucose laboratory data, and body measure examinations (BMI) data from the 2015-2016, 2017-2020 pre-pandemic, and 08/2021-08/2023 NHANES cycles. The 2009-2010, 2010-2012, and 2013-2014 cycles were not used since they had different methodologies for insulin (2009-2010 and 2010-2012 cycles) and glucose (2009-2010, 2010-2012, and 2013-2014 cycles).

```{r}
table_names <- c("DEMO", "DIQ", "GHB", "BMX", "INS", "GLU") # for survey age, diabetes diagnosis age, a1c, bmi, insulin, and glucose

table_names <- append(table_names, c("HIQ", # health insurance
                                     "RXQ_RX")) # prescriptions
                                    

table_suffixes <- c("I", "L") # for 2015-2016, and 2021-2023

table_codes <- map(table_names, paste, table_suffixes, sep = "_")
names(table_codes) <- table_names

nhanes_tables <- list()

for (s in table_suffixes){
  nhanes_tables[[s]] <- list()
  
  table_codes <- purrr::map_vec(table_names, paste, s, sep = "_")
  
  nhanes_tables[[s]] <- map(table_codes, nhanes) %>% # table INS_G will give download length error since INS is in GLU that year
    map(., as_tibble)
  
  names(nhanes_tables[[s]]) <- table_names
}

# pre-pandemic data has prefix instead of suffix, extract separately
p <- map(paste("P", table_names, sep = "_"), nhanes) %>% map(., as_tibble)
names(p) <- table_names

nhanes_tables[["P"]] <- p

# import prescripiton codes
codebook_url <- "https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/1988/DataFiles/RXQ_DRUG.xpt"
temp_file <- tempfile()
download.file(codebook_url, temp_file, mode = "wb", quiet = TRUE)
prescription_codebook <- foreign::read.xport(temp_file) %>% as_tibble()
unlink(temp_file)
```

# Data cleaning

Merge additional data from NHANES. 

```{r message=FALSE, warning=FALSE}

prescription_class <- function(df, prescription_codebook, prescription_category2){
  
  # fill in blank as unspecified
  codebook <- prescription_codebook %>% 
    mutate(RXDDCN1C = case_when(RXDDCN1C == "" ~ "UNSPECIFIED", .default = RXDDCN1C)) # label unspecified prescriptions if first in category
  
  # extract SEQN to merge
  df_summarized <- df %>% distinct(SEQN)
  
  # match prescription to each level 3 category
  codebook_level3 <- codebook %>%
    select(RXDDRGID, 
           RXDDCN1C, RXDDCN2C, RXDDCN3C)
  df <- df %>% 
    left_join(codebook_level3, by = join_by(RXDDRGID))
  
  # loop for each level2 category
  for (cat2 in prescription_category2){

    # find all unique first level 3 categories under defined level 2 category
    cat3 <- codebook %>% 
      filter(RXDDCN1B == cat2) %>% 
      distinct(RXDDCN1C) %>% 
      arrange(desc(RXDDCN1C)) %>%
      pull(RXDDCN1C)
    

    # tidy category names
    tidy_cat2 <- str_replace_all(tolower(cat2), " |-|\\(|\\)", "_")
    
    df_summarized <- df_summarized %>% 
      mutate("{tidy_cat2}" := "No")
    
    # label each prescription to individual level 3 categories under level 2 category
    for (category in cat3){
      # construct tidy column name
      tidy_category <- str_replace_all(tolower(category), " |-|\\(|\\)", "_")
      
      # append level 2 name to unspecified
      tidy_category <- if_else(tidy_category == "unspecified", paste(tidy_category, tidy_cat2, sep = "_"), tidy_category)
      
      df_summarized <- df %>%
        group_by(SEQN) %>%
        # detect level 3 category in any prescription levels
        summarise("{tidy_category}" := as.integer(any(str_detect(RXDDCN1C, category), 
                                                      str_detect(RXDDCN2C, category),
                                                      str_detect(RXDDCN3C, category)))) %>% 
        mutate("{tidy_category}" := case_when(!!sym(tidy_category) == 1 ~ "Yes", .default = "No")) %>%
                 #replace_na(!!sym(tidy_category), 0)) %>% # replace NA from no prescription with 0
        right_join(df_summarized, by = join_by(SEQN))
      
      # change level2 category to positive if any in level3 category is detected
      df_summarized <- df_summarized %>%
        mutate("{tidy_cat2}" := case_when(!!sym(tidy_category) == "Yes" ~ "Yes", .default = !!sym(tidy_cat2)))
      
    }
    
  }

  return(df_summarized)
  
}


# fix colnames to match
# nhanes_tables$I$BPXO <- nhanes_tables$I$BPXO %>%
#   rename_with(., ~stringr::str_replace(.x, "BPX", "BPXO"))

cycle_tables <- list()
for (s in c(table_suffixes, "P")){
  cycle_tables[[s]] <- nhanes_tables[[s]]$DEMO %>% mutate(cycle = s)
  
  for (t in table_names[table_names != "DEMO"]){
    
    if (t == "RXQ_RX" & s != "L"){ # no prescription details in cycle L
      nhanes_tables[[s]][[t]] <- prescription_class(nhanes_tables[[s]][[t]], 
                                                    prescription_codebook, 
                                                    c("ANTIHYPERTENSIVE COMBINATIONS", "ANTIDIABETIC AGENTS"))
        
        # nhanes_tables[[s]][[t]] %>%
        # group_by(SEQN) %>%
        # summarise(insulin = sum(str_detect(RXDDRUG, "INSULIN")),
        #           verapamil = sum(str_detect(RXDDRUG, "VERAPAMIL"))) %>%
        # mutate(insulin = case_when(insulin > 0 ~ 1, .default = 0),
        #        verapamil = case_when(verapamil > 0 ~ 1, .default = 0))
    }
    
    cycle_tables[[s]] <- cycle_tables[[s]] %>% 
      full_join(nhanes_tables[[s]][[t]]) %>% select(-one_of("WTPH2YR")) # remove phebotomy weight, causes duplication issues
    
    
    
  }
  
  # if (s %in% table_suffixes){
  #   cycle_tables[[s]] <- cycle_tables[[s]] %>%
  #     rename(diet_weight = WTDRD1) # diet weight, derived from subset of fasting MEC weight
  #      
  # }
  # 
  # if (s == "P"){
  #   cycle_tables[[s]] <- cycle_tables[[s]] %>%
  #     rename(diet_weight = WTDRD1PP) # diet weight, derived from subset of fasting MEC weight
  # }
  
}
  
# convert factor to numeric to match formatting
cycle_tables$I <- cycle_tables$I %>%
  mutate(DMDHHSIZ = as.numeric(str_extract(DMDHHSIZ, "\\d+")))

merged_nhanes <- bind_rows(cycle_tables)
```

# Export data

Export data for downstream analysis.

```{r}

export_dir <- here::here("exports/NHANES")

if (!dir.exists(export_dir)) {
  dir.create(export_dir, recursive = T)
}

saveRDS(merged_nhanes, file.path(export_dir, "nhanes_raw_data.rds"))

```

# Session information

```{r}

sessionInfo()

```
