---
title: "UAB features joined"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
Although diabetes is broadly classified as type 1 and type 2 diabetes, type 2 diabetes is known to be heterogeneous. Finer classifications of diabetes could aid in the personalization of diabetes treatment. [Ahlqvist et al.](https://doi.org/10.1016/s2213-8587(18)30051-2) recently identified 5 subtypes of diabetes with different risks of diabetic complications using glutamate decarboxylase antibody status, age at diagnoses, body mass index, HbA1c, and homeostatic model assessment 2 estimates of beta-cell function and insulin resistance. 

**Objective:** Join all cleaned records from the UAB Health Systems to feed into clustering algorithms downstream.

# Environment set up

```{r}
library(tidyverse)
```

Set root directory

```{r}
here::i_am("README.md")

```

# Import files
Identify CSV files containing cleaned records from the UAB Health Systems.

```{r}
export_dir <- here::here("exports/UAB/raw")

raw_data_paths <- list.files(export_dir)

## remove export file used to compute homa and previously merged data
remove_index <- grep("homa_cpeptide\\.csv|homa_insulin\\.csv|merged\\.csv", raw_data_paths)
raw_data_paths <- raw_data_paths[-remove_index]

```

Import identified CSV files as a list.

```{r}
paths <- file.path(export_dir, raw_data_paths)

file_names <- c(str_extract(raw_data_paths, "(.*)\\.csv$", group = 1))

# remove suffix in file name for externally calculated homa estimates
file_names <- str_replace_all(file_names, "_homa", "")

records <- lapply(paths, read_csv, show_col_types = F)
names(records) <- file_names
```

# Clean data
Change some columns names to make joining easier and correct formatting errors.

```{r}
# rename columns to distinguish original columns from joined columns
records[["homa_cpeptide"]] <- records[["homa_cpeptide"]] %>% rename(homa_cpeptide_cpeptide = cpeptide,
                                                                              homa_cpeptide_cpeptide_units = cpeptide_units,
                                                                              homa_cpeptide_glucose  = glucose,
                                                                              homa_cpeptide_glucose_units = glucose_units)

# correct date formatting messed up by importing into Excel to calculate HOMA
records[["homa_cpeptide"]] <- records[["homa_cpeptide"]] %>% mutate(homa_cpeptide_date = mdy(homa_cpeptide_date))

# rename columns to distinguish original columns from joined columns
records[["homa_insulin"]] <- records[["homa_insulin"]] %>% rename(homa_insulin_insulin = insulin,
                                                                            homa_insulin_insulin_units = insulin_units,
                                                                            homa_insulin_glucose  = glucose,
                                                                            homa_insulin_glucose_units = glucose_units)

# correct date formatting messed up by importing into Excel to calculate HOMA
records[["homa_insulin"]] <- records[["homa_insulin"]] %>% mutate(homa_insulin_date = mdy(homa_insulin_date))

# rename columns to match file name
colnames(records[["triglyceride"]]) <- str_replace_all(colnames(records[["triglyceride"]]), "tg", "triglyceride")

# rename columns in u_ketone positive to distinguish from u_ketone 
colnames(records[["u_ketone_first_positive"]]) <- str_replace_all(colnames(records[["u_ketone_first_positive"]]), "u_ketone", "u_ketone_first_positive")

# convert u_ketone_positive from date time to date only
records[["u_ketone_first_positive"]] <- records[["u_ketone_first_positive"]] %>%
  mutate(u_ketone_first_positive_date_time = date(u_ketone_first_positive_date_time)) %>%
  rename(u_ketone_first_positive_date = u_ketone_first_positive_date_time)

```

Source the cleaning function `keep_first`. `keep_first` finds and returns the first lab test of the day for each patients and discards the rest.

```{r}
source("/data/user/brianlu/projects/precision_diabetes/Clustering/functions_feature_cleaning.R")
```

Find and keep the first glucose, insulin, and cpeptide test of the day for each patient.

```{r}
records[["glucose"]] <- keep_first(records[["glucose"]])
records[["insulin"]] <- keep_first(records[["insulin"]])
records[["cpeptide"]] <- keep_first(records[["cpeptide"]])
```

Calculate age at first diabetes diagnoses in the records for each patient.

```{r}
joined_records <- inner_join(records[["diagnoses"]], records[["birthdays"]], by = join_by(DE_ID)) %>%
  mutate(diabetes_age = as.numeric(diabetes_date - birth_date, "days")/365)
```

# Join records
Join demographic data to diagnoses and encounters.

```{r}
joined_records <- inner_join(records[["patient_demo"]], joined_records, by = join_by(DE_ID))
```

Left join all records on diagnoses. Only the test result closest to the initial diabetes diagnoses date for each patient was kept.

```{r}
# find index position of all test results
index = seq(1, length(records))[-grep("diagnoses|birthdays|demo", names(records))]

for (i in index){
  # left join on diagnoses and birthdays
  joined_records <- left_join(joined_records, records[[i]], by = join_by(DE_ID), relationship = "many-to-many")
  
  # calculate days to diagnoses and keep only nearest test results
  test_name <- names(records)[i]
  date_name <- paste(test_name, "date", sep = "_")
  test_days <- paste(test_name, "diabetes_days", sep = "_")
  joined_records <- mutate(.data = joined_records, "{test_days}" := as.numeric(.data[[date_name]] - diabetes_date, "days")) %>%
    group_by(DE_ID) %>%
    filter(abs(.data[[test_days]]) == min(abs(.data[[test_days]])) | 
             is.na(.data[[test_days]])) %>% # keep the test with closest to the diagnoses date
    filter(.data[[test_days]] == max(.data[[test_days]]) | 
             is.na(.data[[test_days]])) %>% # if test is equal in days ahead and after the diagnoses date, keep the one after
    distinct()
    
}

```

# Export results
Export joined records as CSV file.

```{r}
# export
write_csv(joined_records, file.path(export_dir, "merged.csv"))
```

# Session information

```{r}
sessionInfo()
```
