---
title: "UAB diagnoses"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
Although diabetes is broadly classified as type 1 and type 2 diabetes, type 2 diabetes is known to be heterogeneous. Finer classifications of diabetes could aid in the personalization of diabetes treatment. 

**Objective:** Extract and clean the diagnoses records from the UAB Health Systems.

# Environment set up

```{r}
library(RSQLite)
library(tidyverse)
```

Set root directory

```{r}
here::i_am("README.md")
```

# Feature extraction
Diabetes records from 2020 to 2024 were downloaded from the University of Alabama at Birmingham's (UAB) Health System by Andrew Crouse onto Microsoft Teams. The records were uploaded to Cheaha and imported into the `crouse_2024-06-17.db` SQL database using SQLite. 

The required records were retrieved from `crouse_2024-06-17.db` as separate tibbles. 

```{r}
# connect to SQL database
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2024-06-17.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

diagnoses <- list()
```

Diabetes diagnoses were retrieved by codes:

* ICD-9-CM:
  + 249 
  + 250
* ICD-10-CM
  + E08
  + E09
  + E10
  + E11
  + E13

Gestational diabetes diagnoses were not included.

```{r}
# pull required information from database
## diabetes diagnoses
diagnoses$diabetes <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                        FROM diagnoses
                        WHERE (VOCABULARY = "ICD-10-CM" 
                            and (CODE_VALUE LIKE "E08%"
                              or CODE_VALUE LIKE "E09%"
                              or CODE_VALUE LIKE "E10%"
                              or CODE_VALUE LIKE "E11%"
                              or CODE_VALUE LIKE "E13%"))
                          or (VOCABULARY = "ICD-9-CM" 
                            and (CODE_VALUE LIKE "250%"
                              or CODE_VALUE LIKE "249%"))') %>% as_tibble()

```

Diabetic neuropathy diagnoses were retrieved by codes:

* ICD-9-CM:
  + 249.6
  + 250.6
* ICD-10-CM:
  + E08.4
  + E09.4
  + E10.4
  + E13.4 

```{r}
diagnoses$neuropathy <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                         FROM diagnoses
                         WHERE (VOCABULARY = "ICD-10-CM"
                            and (CODE_VALUE LIKE "E08.4%"
                              or CODE_VALUE LIKE "E09.4%"
                              or CODE_VALUE LIKE "E10.4%"
                              or CODE_VALUE LIKE "E11.4%"
                              or CODE_VALUE LIKE "E13.4%"))
                          or (VOCABULARY = "ICD-9-CM"
                            and (CODE_VALUE LIKE "250.6%"
                              or CODE_VALUE LIKE "249.6%"))') %>% as_tibble()

```

Diabetic retinopathy diagnoses were retrieved by codes:

* ICD-9-CM:
  + 362.0
  + 249.5 with "retinopathy" in code description
  + 250.5 with "retinopathy" in code description 
* ICD-10-CM
  + E08.31 through E08.35, and E08.37
  + E09.31 through E09.35, and E09.37
  + E10.31 through E10.35, and E10.37
  + E11.31 through E11.35, and E11.37
  + E13.31 through E13.35, and E13.37
  

```{r}
## used GLOB with * wildcard instead of LIKE and % for ranges, special to SQLite3
diagnoses$retinopathy <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                        FROM diagnoses
                        WHERE (VOCABULARY = "ICD-10-CM"
                            and (CODE_VALUE GLOB "E08.3[123457]*"
                              or CODE_VALUE GLOB "E09.3[123457]*" 
                              or CODE_VALUE GLOB "E10.3[123457]*" 
                              or CODE_VALUE GLOB "E11.3[123457]*"
                              or CODE_VALUE GLOB "E13.3[123457]*"))
                          or (VOCABULARY = "ICD-9-CM"
                            and (CODE_VALUE LIKE "362.0%"
                              or (CODE_VALUE LIKE "250.5%"
                                and CODE_DESCRIPTION LIKE "%retinopathy%")
                              or (CODE_VALUE LIKE "249.5%"
                                and CODE_DESCRIPTION LIKE "%retinopathy%")))') %>% as_tibble()

  
```


End stage renal disease (ESRD) diagnoses were retrieved by codes:

* ICD-9-CM:
  + 585.5
  + 585.6
* ICD-10-CM:
  + N18.5
  + N18.6


```{r}
## end stage renal disease
diagnoses$esrd <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                        FROM diagnoses
                        WHERE (VOCABULARY = "ICD-9-CM" and 
                            (CODE_VALUE LIKE "585.5%" or CODE_VALUE LIKE "585.6%"))
                          or (VOCABULARY = "ICD-10-CM" and 
                            (CODE_VALUE LIKE "N18.5%" or CODE_VALUE LIKE "N18.6%"))') %>% as_tibble()

```

Nephropathy diagnoses were retrieved by codes:

* ICD-9-CM: 
  + 249.4
  + 250.4
* ICD-10-CM:
  + E08.2
  + E09.2
  + E10.2
  + E11.2
  + E13.2
  
Nephropathy also included ESRD diagnoses.
  
```{r}
diagnoses$nephropathy <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                         FROM diagnoses
                         WHERE (VOCABULARY = "ICD-10-CM"
                            and (CODE_VALUE LIKE "E08.2"
                              or CODE_VALUE LIKE "E09.2%"
                              or CODE_VALUE LIKE "E10.2%"
                              or CODE_VALUE LIKE "E11.2%"
                              or CODE_VALUE LIKE "E13.2%"
                              or CODE_VALUE LIKE "N18%"))
                          or (VOCABULARY = "ICD-9-CM"
                            and (CODE_VALUE LIKE "250.4%"
                              or CODE_VALUE LIKE "249.4%"
                              or CODE_VALUE LIKE "585%"))') %>% as_tibble()

# also include ESRD
diagnoses$nephropathy <- full_join(diagnoses$nephropathy, diagnoses$esrd)
```


Non-alcoholic fatty liver disease (NAFLD) and non-alcoholic steatohepatitis (NASH) diagnoses were retrieved by codes:

* ICD-9-CM:
  + 571.8 
* ICD-10-CM:
  + K76.0
  + K75.81

Hepatosplenomegaly diagnoses were removed.

```{r}
## non-alcoholic fatty liver disease or nonalcoholic steatohepatitis 
diagnoses$nafld <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                        FROM diagnoses
                        WHERE (VOCABULARY = "ICD-9-CM" and CODE_VALUE = "571.8")
                          or (VOCABULARY = "ICD-10-CM" and (CODE_VALUE = "K76.0"
                            or CODE_VALUE = "K75.81"))') %>% as_tibble()

diagnoses$nafld <- diagnoses$nafld %>% filter(!grepl("hepatosplenomegaly", .$CODE_DESCRIPTION, ignore.case = T)) # enlarged liver
```

Cardiovascular disease was defined as myocardial infarction or stroke, as used in [Fedotkina et al.](https://doi.org/10.3389/fgene.2021.637945), using codes:

* ICD-9-CM: 
  + 410.0 through 410.5, 410.8, and 410.9
  + 411.0 and 411.8
  + 431
  + 432
  + 433 with code suffix ending in 1
  + 434 with code suffix ending in 1
* ICD-10-CM: 
  + I21
  + I24
  + I61 through I64

```{r}
## cardiovascular disease
diagnoses$cvd <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                        FROM diagnoses
                        WHERE (VOCABULARY = "ICD-9-CM" 
                            and (CODE_VALUE GLOB "410.[01234589]*"
                              or CODE_VALUE GLOB "411.[08]*"
                              or CODE_VALUE LIKE "431%"
                              or CODE_VALUE LIKE "432%"
                              or CODE_VALUE LIKE "433.%1"
                              or CODE_VALUE LIKE "434.%1"))
                          or (VOCABULARY = "ICD-10-CM"
                            and (CODE_VALUE LIKE "I21%"
                              or CODE_VALUE LIKE "I24%"
                              or CODE_VALUE LIKE "I61%"
                              or CODE_VALUE LIKE "I62%"
                              or CODE_VALUE LIKE "I63%"
                              or CODE_VALUE LIKE "I64%"))') %>% as_tibble()

diagnoses$cvd <- diagnoses$cvd %>% filter(!grepl("post-traumatic", .$CODE_DESCRIPTION, ignore.case = T))
```

Cardiovascular disease was further separated into myocardial infarction and stroke. 

```{r}
## myocardial infarction
diagnoses$mi <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                        FROM diagnoses
                        WHERE (VOCABULARY = "ICD-9-CM" 
                            and (CODE_VALUE GLOB "410.[01234589]*"
                              or CODE_VALUE GLOB "411.[08]*"))
                          or (VOCABULARY = "ICD-10-CM"
                            and (CODE_VALUE LIKE "I21%"
                              or CODE_VALUE LIKE "I24%"))') %>% as_tibble()

## stroke
diagnoses$stroke <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                        FROM diagnoses
                        WHERE (VOCABULARY = "ICD-9-CM" 
                            and (CODE_VALUE LIKE "431%"
                              or CODE_VALUE LIKE "432%"
                              or CODE_VALUE LIKE "433.%1"
                              or CODE_VALUE LIKE "434.%1"))
                          or (VOCABULARY = "ICD-10-CM"
                            and (CODE_VALUE LIKE "I61%"
                              or CODE_VALUE LIKE "I62%"
                              or CODE_VALUE LIKE "I63%"
                              or CODE_VALUE LIKE "I64%"))') %>% as_tibble()
```

Diabetic ketoacidosis was defined by codes:

* ICD-9-CM:
  + 249.1
  + 250.1
* ICD-10-CM:
  + E08.1
  + E09.1
  + E10.1
  + E11.1
  + E13.1

```{r}
diagnoses$dka <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                         FROM diagnoses
                         WHERE (VOCABULARY = "ICD-9-CM" 
                              and (CODE_VALUE = "249.1"
                                or CODE_VALUE LIKE "249.1%"
                                or CODE_VALUE = "250.1"
                                or CODE_VALUE LIKE "250.1%"))
                            or (VOCABULARY = "ICD-10-CM" 
                              and (CODE_VALUE = "E08.1" 
                              or CODE_VALUE LIKE "E08.1%"
                              or CODE_VALUE = "E09.1"
                              or CODE_VALUE LIKE "E09.1%"
                              or CODE_VALUE = "E10.1" 
                              or CODE_VALUE LIKE "E10.1%"
                              or CODE_VALUE = "E11.1"
                              or CODE_VALUE LIKE "E11.1%"
                              or CODE_VALUE = "E13.1"
                              or CODE_VALUE LIKE "E13.1%"))') %>% as_tibble()

```

COVID-19 was defined by codes:

* U07.1 for COVID-19 after April 1, 2020
* U09 for post-COVID-19 conditions
* B97.29 for COVID-19 before April 1, 2020
* Z86.16 for personal history of COVID-19, effective January 1, 2021

Codes not included since no diagnoses were found:

* U07.2 for COVID-19, virus not identified
* U08 for personal history of COVID-19
* U10 for multisystem inflammatory syndrome associated with COVID-19

Diagnoses found before the first confirmed case of COVID-19 in the United States on January 20, 2020, as reported by the CDC, were removed.

```{r}
diagnoses$covid19 <- dbGetQuery(conn, 'SELECT DE_ID, DE_ID_VISIT_NUMBER, DIAG_DT_TM, DIAG_TYPE, VOCABULARY, CODE_VALUE, CODE_DESCRIPTION 
                         FROM diagnoses
                         WHERE VOCABULARY = "ICD-10-CM"
                              and (CODE_VALUE LIKE "U07.1%"
                                or CODE_VALUE LIKE "U09%"
                                or CODE_VALUE = "Z86.16"
                                or CODE_VALUE = "B97.29")') %>% as_tibble() %>%
  mutate(date = date(mdy_hms(DIAG_DT_TM))) %>%
  filter(date >= mdy("01-20-2020")) %>%
  select(-date)
```


Encounters were used to match patient visits across different tibbles.

```{r}
## patient encounters
encounters <- dbGetQuery(conn, "SELECT DE_ID, DE_ID_VISIT_NUMBER, ENCOUNTER_TYPE_CLASS, ENCOUNTER_STATUS, ENCOUNTER_DATE, DISCHARGE_DATE FROM encounters") %>% as_tibble()
```

Close SQL connection.

```{r}
dbDisconnect(conn)
```

# Data cleaning

## Encounters
Each patient visit was tracked as an encounter with a unique visit number. 
```{r}
# convert dates from string to datetime
encounters <- encounters %>% mutate(encounter_date = date(mdy_hms(ENCOUNTER_DATE)),
                                  discharge_date = date(mdy_hms(DISCHARGE_DATE))) %>%
  select(-c(ENCOUNTER_DATE, DISCHARGE_DATE))

# remove stays with negative values
encounters <- encounters %>% mutate(stay_days = as.numeric(discharge_date - encounter_date, "days"))
encounters_filtered <- encounters %>% filter(stay_days >= 0 | is.na(discharge_date))

# remove encounters with no type
encounters_filtered <- encounters_filtered %>% filter(ENCOUNTER_TYPE_CLASS != "")

# remove duplicates
encounters_cleaned <- encounters_filtered %>% distinct(DE_ID, DE_ID_VISIT_NUMBER, ENCOUNTER_TYPE_CLASS, .keep_all = T)
```

## Diagnoses
Since UAB records did not indicate whether a diagnosis was the patient's first diagnosis, the earliest diagnosis in the records for a patient was assumed to be the patient's first diagnosis. 

A custom function was written to 

1. remove duplicates, working diagnoses, and diagnoses without a matching record in `encounters`
2. rename columns to keep track of diagnoses after joining

```{r}
first_diagnosis <- function(df, diag, encounter = NULL){
  # remove working diagnoses
  df_filtered <- df %>% filter(DIAG_TYPE != "Working") %>%
    distinct()
  
  # matched encounters if available
  if (!is.null(encounter)){
    df_filtered <- df_filtered %>% 
      left_join(encounter, by = c("DE_ID", "DE_ID_VISIT_NUMBER")) %>%
      rename("{diag}_encounter_type_class" := ENCOUNTER_TYPE_CLASS,
             "{diag}_encounter_status" := ENCOUNTER_STATUS,
             "{diag}_encounter_date" := encounter_date, 
             "{diag}_discharge_date" := discharge_date,
             "{diag}_stay_days" := stay_days)
    
  }
  
  # store diagnosis date as separate variable
  diag_date = paste0(diag, "_date")
  
  # rename diagnosis date
  df_filtered <- df_filtered %>%
    mutate("{diag_date}" := date(mdy_hms(DIAG_DT_TM))) %>%
    relocate({{diag_date}}, .after = DE_ID_VISIT_NUMBER) %>%
    select(-DIAG_DT_TM)
  
  # find earliest diagnosis date for each DE_ID
  df_initial <- df_filtered %>%
    group_by(DE_ID) %>% 
    arrange(across(all_of(diag_date)), .by_group = T) %>% # arrange by ascending order within group
    distinct(DE_ID, .keep_all = T) %>% # take first row of each group, ie earliest date
    ungroup()
  
  # rename other variables
  df_cleaned <- df_initial %>%
    rename("{diag}_visit_number" := DE_ID_VISIT_NUMBER,
         "{diag}_diag_type" := DIAG_TYPE,
         "{diag}_vocabulary" := VOCABULARY,
         "{diag}_code" := CODE_VALUE,
         "{diag}_description" := CODE_DESCRIPTION)
  

  return(df_cleaned)
}
```

Loop function through list of diagnoses to clean diagnoses.

```{r}
diagnoses_cleaned <- list()

for (n in names(diagnoses)){
  diagnoses_cleaned[[n]] <- first_diagnosis(diagnoses[[n]], n, encounters_cleaned)
}
```

Remove patients diagnosed with diabetes before 2020-01-01.

```{r}
diagnoses_cleaned$diabetes <- diagnoses_cleaned$diabetes %>% filter(diabetes_date >= "2020-01-01")
```

To make absolutely sure no patients in this cohort was already in the 2010-2019 cohort, filter using MRN.

```{r}
# connect to SQL database and extract diabetes diagnoses from pre-pandemic cohort
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2023-06-26.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)
cohort_2010 <- dbGetQuery(conn, 'SELECT DISTINCT DE_ID
                        FROM diagnoses
                        WHERE (VOCABULARY = "ICD-10-CM" 
                            and (CODE_VALUE LIKE "E08%"
                              or CODE_VALUE LIKE "E09%"
                              or CODE_VALUE LIKE "E10%"
                              or CODE_VALUE LIKE "E11%"
                              or CODE_VALUE LIKE "E13%"))
                          or (VOCABULARY = "ICD-9-CM" 
                            and (CODE_VALUE LIKE "250%"
                              or CODE_VALUE LIKE "249%"))') %>% as_tibble()

# connect to SQL database for pre-pandemic cohort MRN
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2024-01-10.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

# get matching MRNS for DE_ID for pre-pandemic
mrn_2010 <- dbGetQuery(conn, "SELECT MRNS, DE_ID FROM mrns") %>% 
  as_tibble() %>% 
  filter(DE_ID %in% cohort_2010$DE_ID)

# connect to SQL database for post-pandemic cohort MRN
database_path <- "/data/user/brianlu/projects/precision_diabetes/SQL_database/crouse_2024-06-17.db"
conn <- dbConnect(RSQLite::SQLite(), database_path)

# get matching MRNS for DE_ID for post-pandemic
mrn_2020 <- dbGetQuery(conn, "SELECT MRN, DE_ID FROM mrns") %>% 
  as_tibble() %>% 
  filter(DE_ID %in% diagnoses_cleaned$diabetes$DE_ID)

# find MRNS from post-pandemic that's also in the pre-pandemic cohort
duplicate_mrn <- mrn_2020 %>% 
  semi_join(mrn_2010, by = c("MRN" = "MRNS"))

# convert MRN back to DE_ID, since single MRN can match to multiple DE_ID, and find unique DE_ID
unique_deid <- mrn_2020 %>% 
  anti_join(duplicate_mrn, by = join_by(DE_ID))

# keep only unique DE_ID
diagnoses_cleaned$diabetes <- diagnoses_cleaned$diabetes %>%
  filter(DE_ID %in% unique_deid$DE_ID)

```



Label type of diabetes diagnoses.

```{r}

# construct ICD-9 type 1 and type 2 codes
icd9_diabetes <- paste("^250", seq(0, 9), sep = "\\.")
icd9_type1 <- paste(c(paste0(icd9_diabetes, 1), paste0(icd9_diabetes, 3)), collapse = "|")
icd9_type2 <- paste(c(paste0(icd9_diabetes, 0), paste0(icd9_diabetes, 2)), collapse = "|")

# construct typoe 1 and type 2 diabetes codes
type1_codes <- paste("^E10", icd9_type1, sep = "|")
type2_codes <- paste("^E11", icd9_type2, sep = "|")

diagnoses_cleaned$diabetes <- diagnoses_cleaned$diabetes %>%
  mutate(diabetes_type = 
           # by code description first
           ifelse(grepl("latent autoimmune|LADA|1\\.5", .$diabetes_description, ignore.case = T), "LADA",
                  ifelse(grepl("maturity|mody", .$diabetes_description, ignore.case = T), "MODY",
                         ifelse(grepl("type 2|type II|NIDDM|noninsulin dependent|non-insulin dependent", 
                                      .$diabetes_description, ignore.case = T), "Type 2",
                                ifelse(grepl("type 1|type I|IDDM|insulin dependent", .$diabetes_description, ignore.case = T), "Type 1",
                                       ifelse(grepl("gestational", .$diabetes_description, ignore.case = T), "Gestational",
                                              ifelse(grepl("Other pre-existing|unspecified pre-existing|unspecified diabetes", 
                                                    .$diabetes_description, ignore.case = T), "Unspecified",
            # then by code                                  
            ifelse(grepl("^249|^E08|^E09", .$diabetes_code), "Secondary",
                   ifelse(grepl(type2_codes, .$diabetes_code), "Type 2",
                          ifelse(grepl(type1_codes, .$diabetes_code), "Type 1",
                                 ifelse(grepl("^E13", .$diabetes_code), "Other",
                                        ifelse(grepl("^O24|^648", .$diabetes_code), "Gestational",
                                               "Unspecified"))))))))))))

# Find MODY, since MODY can be misdiagnosed as type 2 on initial diagnoses
mody_de_id <- diagnoses$diabetes %>% 
  filter(grepl("maturity|mody", .$CODE_DESCRIPTION, ignore.case = T)) %>% 
  distinct(DE_ID) %>% 
  pull()

# relabel MODY diagnoses
diagnoses_cleaned$diabetes$diabetes_type <- ifelse(diagnoses_cleaned$diabetes$DE_ID %in% mody_de_id, 
                                     "MODY", 
                                     diagnoses_cleaned$diabetes$diabetes_type)

# should not have any gestational diabetes
table(diagnoses_cleaned$diabetes$diabetes_type)
```

Join all diagnoses on diabetes diagnoses. Time to first diabetes diagnosis was calculated in days.

```{r}

combined_diagnoses <- diagnoses_cleaned$diabetes

for (i in 2:length(diagnoses_cleaned)){
  diagnosis_name <- names(diagnoses_cleaned)[i]
  diagnosis_date <- paste(names(diagnoses_cleaned)[i], "date", sep = "_")
  combined_diagnoses <- left_join(combined_diagnoses, diagnoses_cleaned[[i]], by = join_by(DE_ID)) %>%
    mutate("{diagnosis_name}_diabetes_days" := as.numeric(.[[diagnosis_date]] - diabetes_date, "days"), 
      "{diagnosis_name}" := ifelse(is.na(.[[diagnosis_date]]), 0, 1))
}
```

# Export results

```{r}
# create export directory if doesn't exist
export_dir <- here::here("exports/UAB/raw")

if (!dir.exists(export_dir)) {
  dir.create(export_dir, recursive = T)
}

write_csv(combined_diagnoses, file.path(export_dir, "diagnoses.csv"))
```

# Session information

```{r}
sessionInfo()
```

