---
title: "HOMA scaling models"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

**Objective:** Prepare fixed-effects only models of the trained HOMA scaling models for release.

# Environment set up

```{r}
library(tidyverse)
library(lme4)
library(lmerTest)
```

Set root directory

```{r}
here::i_am("README.md")

export_dir <- here::here("exports/HOMA_regression")
```


# Generate public release models

Create custom function to use on lists.

```{r}

predict_release <- function(object, newdata) {

  X <- model.matrix(
    object$terms,
    newdata,
    xlev = object$xlevels,
    contrasts.arg = object$contrasts
  )

  drop(X %*% object$beta)
}

```


Strip models down to minimum required fixed effects for public release.

```{r}
# HOMA1-B model
homa1b_model <- readRDS(here::here("exports/HOMA_regression/homa1b_model.RDS"))

homa1b_fe_model <- list(
  beta = fixef(homa1b_model),
  terms = delete.response(terms(homa1b_model)),
  xlevels = .getXlevels(terms(homa1b_model), model.frame(homa1b_model)),
  contrasts = attr(model.matrix(homa1b_model), "contrasts"),
  predict = predict_release
)

# HOMA1-IR model
homa1ir_model <- readRDS(here::here("exports/HOMA_regression/homa1ir_model.RDS"))

homa1ir_fe_model <- list(
  beta = fixef(homa1ir_model),
  terms = delete.response(terms(homa1ir_model)),
  xlevels = .getXlevels(terms(homa1ir_model), model.frame(homa1ir_model)),
  contrasts = attr(model.matrix(homa1ir_model), "contrasts"),
  predict = predict_release
)

```


# Test release models

Read in test data.

```{r}

test_data <- read_csv(here::here("exports/HOMA_regression/homa_regression_test_data.csv"), show_col_types = F)

```

The new predicted HOMA1-B values are identical to those predicted by the original model.

```{r}
test_data <- test_data %>%
     # re.form = NA to specify no random effects
     mutate(homa1b_pred = exp(predict(homa1b_model, newdata = test_data, re.form = NA)),
            homa1b_release_pred = exp(homa1b_fe_model$predict(homa1b_fe_model, newdata = test_data)))

identical(test_data$homa1b_pred, test_data$homa1b_release_pred)

```

The new predicted HOMA1-IR values are identical to those predicted by the original model.

```{r}
test_data <- test_data %>%
     # re.form = NA to specify no random effects
     mutate(homa1ir_pred = exp(predict(homa1ir_model, newdata = test_data, re.form = NA)),
            homa1ir_release_pred = exp(homa1ir_fe_model$predict(homa1ir_fe_model, newdata = test_data)))

identical(test_data$homa1ir_pred, test_data$homa1ir_release_pred)

```

# Export minimal release models

```{r}

saveRDS(homa1b_fe_model, here::here(export_dir, "homa1b_fe_model.RDS"))
saveRDS(homa1ir_fe_model, here::here(export_dir, "homa1ir_fe_model.RDS"))

```

# Session information

```{r}
sessionInfo()
```