---
title: "HOMA regression cluster testing"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
HOMA1 estimates calculated with insulin and c-peptide may results in differences. 

**Objective:** Extract additional clustering features from NHANES and UAB data for the HOMA regression test data to use in cluster assignment.

# Environment setup

```{r}
library(tidyverse)
library(nhanesA)
```

# Data extraction

```{r}
table_names <- c("L10", "BMX") # for a1c, bmi

table_suffixes <- c("nosuffix", "B", "C") # for 1999-2000, 2001-2002, 2003-2004

nhanes_tables <- list()

for (s in table_suffixes){
  nhanes_tables[[s]] <- list()
  
  table_codes <- purrr::map_vec(table_names, paste, s, sep = "_")
  table_codes <- map(table_codes, ~str_remove(string = .x, pattern = "_nosuffix$"))
  
  nhanes_tables[[s]] <- map(table_codes, nhanes) %>% # table INS_G will give download length error since INS is in GLU that year
    map(., as_tibble)
  
  names(nhanes_tables[[s]]) <- table_names
}

# 1999-2000 cycle HbA1c is under LAB10 instead of L10
nhanes_tables[["nosuffix"]][["L10"]] <- nhanes("LAB10") %>% as_tibble()
```

Merge cycles.

```{r}
cycle_tables <- list()

for (s in table_suffixes){
  cycle_tables[[s]] <- nhanes_tables[[s]]$L10 %>% mutate(cycle = s)
  
  for (t in table_names[table_names != "L10"]){
    
    cycle_tables[[s]] <- cycle_tables[[s]] %>% 
      full_join(nhanes_tables[[s]][[t]]) #%>% select(-one_of("WTPH2YR")) # remove phebotomy weight, causes duplication issues
  }
}

merged_nhanes <- bind_rows(cycle_tables)
```

# Data cleaning

Import test data from HOMA regression.

```{r}
export_dir <- here::here("exports", "HOMA_regression")

test_data <- read_csv(file.path(export_dir, "homa_regression_test_data.csv"), show_col_types = F)

nhanes_test_data <- test_data %>% filter(source == "nhanes")

```

Append additional data from NHANES.

```{r}
nhanes_test_data <- left_join(nhanes_test_data, merged_nhanes, by = c("DE_ID" = "SEQN"))
```


Filter to participants with diabetes.

```{r}
nhanes_test_data <- nhanes_test_data %>% 
  filter(source == "nhanes" & (DIQ010 == "Yes" | glucose >= 7 | LBXGH >= 6.5)) %>%
  # filter to HOMA range
  filter(glucose > 3.5 & glucose <= 25 &
           cpeptide >= 0.2 & cpeptide <= 3.5 & 
           insulin >= 20 & insulin <= 400) %>%
  select(DE_ID, source, LBXGH, BMXBMI, RIDAGEYR, 
         cpeptide_homa1b, cpeptide_homa1ir, insulin_homa1b, insulin_homa1ir, 
         homa1b_pred, homa1ir_pred) %>%
  drop_na() %>%
  rename(a1c = LBXGH,
         bmi = BMXBMI,
         diabetes_age = RIDAGEYR,
         homa1_cpeptide_b = cpeptide_homa1b, 
         homa1_cpeptide_ir = cpeptide_homa1ir, 
         homa1_insulin_b = insulin_homa1b, 
         homa1_insulin_ir = insulin_homa1ir,
         homa1_pred_b = homa1b_pred,
         homa1_pred_ir = homa1ir_pred) %>%
  # impute gad as negative
  mutate(gad = 0)
```

Import UAB post-pandemic data to extract features needed for cluster assignment.

```{r}
# import additional clustering features
uab_post <- read_csv(here::here("exports/UAB/uab_mapped_outpatient.csv"), show_col_types = F) %>%
  # filter to records with insulin, c-peptide, and glucose tested at the same time
  filter(homa_cpeptide_date == homa_insulin_date & 
           homa_cpeptide_glucose == homa_insulin_glucose) %>%
  select(DE_ID, gad, a1c, bmi, diabetes_age) 
  
uab_pre <- read_csv(file.path("/data/user/brianlu/projects/precision_diabetes/Clustering/exports", 
                              "manuscript_Ahlqvist_analysis_2024-05-03/outpatient.csv"), show_col_types = F) %>%
  # filter to records with insulin, c-peptide, and glucose tested at the same time
  filter(homa_cpeptide_date == homa_insulin_date & 
           homa_cpeptide_glucose == homa_insulin_glucose) %>%
  select(DE_ID, gad, a1c, bmi, diabetes_age, 
         # also get raw values and known cluster
         homa_cpeptide_cpeptide, homa_cpeptide_glucose, homa_insulin_insulin, homa_insulin_glucose, Ahlqvist)

uab_data <- bind_rows(list(uab_pre, uab_post))

uab_test_data <- test_data %>% 
  filter(source == "uab") %>%
  # filter to HOMA range
  filter(glucose > 3.5 & glucose <= 25 &
           cpeptide >= 0.2 & cpeptide <= 3.5 & 
           insulin >= 20 & insulin <= 400) %>%
  inner_join(uab_data, by = join_by(DE_ID)) %>%
  select(DE_ID, source, gender, race, 
         gad, a1c, bmi, diabetes_age, 
         cpeptide_homa1b, cpeptide_homa1ir, insulin_homa1b, insulin_homa1ir, 
         homa1b_pred, homa1ir_pred,
         homa_cpeptide_cpeptide, homa_cpeptide_glucose, homa_insulin_insulin, homa_insulin_glucose, Ahlqvist) %>%
  rename(homa1_cpeptide_b = cpeptide_homa1b, 
         homa1_cpeptide_ir = cpeptide_homa1ir, 
         homa1_insulin_b = insulin_homa1b, 
         homa1_insulin_ir = insulin_homa1ir,
         homa1_pred_b = homa1b_pred,
         homa1_pred_ir = homa1ir_pred)

```


```{r}

homa1b_model <- readRDS(here::here("exports/HOMA_regression/homa1b_model.RDS"))
homa1ir_model <- readRDS(here::here("exports/HOMA_regression/homa1ir_model.RDS"))

# calculate HOMA1
uab_test_data <- uab_test_data %>%
  mutate(cluster_homa1_cpeptide_b = (20 * 6 * homa_cpeptide_cpeptide) / (homa_cpeptide_glucose - 3.5),
         cluster_homa1_cpeptide_ir = (homa_cpeptide_glucose * 6 * homa_cpeptide_cpeptide) / 22.5,
         insulin_homa1b = (20 * homa_insulin_insulin/6) / (homa_insulin_glucose - 3.5),
         insulin_homa1ir = (homa_insulin_glucose * homa_insulin_insulin/6) / 22.5) 



uab_test_data <- uab_test_data %>%
  # re.form = NA to specify no random effects
  mutate(cluster_homa1b_pred = exp(predict(homa1b_model, newdata = uab_test_data, re.form = NA)), 
         cluster_homa1ir_pred = exp(predict(homa1ir_model, newdata = uab_test_data, re.form = NA))) %>%
  # rename clustering insulin HOMA1 to keep separate
  rename(cluster_homa1_insulin_b = insulin_homa1b,
         cluster_homa1_insulin_ir = insulin_homa1ir)

```

Merge and export data for cluster assignment.

```{r}
cleaned_test_data <- bind_rows(list(uab_test_data, nhanes_test_data))

write_csv(cleaned_test_data, here::here("exports", "HOMA_regression/cluster_assignment_test_data.csv"))
```

# Session information

```{r}
sessionInfo()
```

