---
title: "HOMA1 regression"
author: "Brian Lu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
HOMA1 estimates calculated with insulin and c-peptide may results in differences. 

**Objective:** Use linear regression to convert between c-peptide HOMA1 estimates and insulin HOMA1 estimates.

# Environment set up

```{r}
library(tidyverse)
library(nhanesA)

library(gridExtra)

library(lme4)
library(lmerTest)
library(MuMIn)

set.seed(42)
```

Set root directory

```{r}
here::i_am("README.md")
```

# Data cleaning

## UAB data

Import all available records for calculating HOMA1 estimates between 2010-2019.

```{r}

import_dir <- "/data/user/brianlu/projects/precision_diabetes/Clustering/exports/features_Ahlqvist_2024-02-05"
glucose2010 <- read_csv(file.path(import_dir, "glucose.csv"), show_col_types = F)
cpeptide2010 <- read_csv(file.path(import_dir, "cpeptide.csv"), show_col_types = F)
insulin2010 <- read_csv(file.path(import_dir, "insulin.csv"), show_col_types = F)
birthday2010 <- read_csv(file.path(import_dir, "birthdays.csv"), show_col_types = F)

demo_dir <- "/data/user/brianlu/projects/precision_diabetes/Clustering/exports/features_additional_2024-02-05"
demo2010 <- read_csv(file.path(demo_dir, "patient_demo.csv"), show_col_types = F)
```

Filter data to only records containing glucose, insulin and c-peptide values measured at the same time.

```{r}
data2010 <- glucose2010 %>%
  inner_join(cpeptide2010, by = c("DE_ID" = "DE_ID",
                              "glucose_visit_number" = "cpeptide_visit_number", 
                              "glucose_date_time" = "cpeptide_date_time")) %>%
  inner_join(insulin2010, by = c("DE_ID" = "DE_ID",
                              "glucose_visit_number" = "insulin_visit_number", 
                              "glucose_date_time" = "insulin_date_time")) %>%
  inner_join(birthday2010, by = join_by(DE_ID)) %>%
  inner_join(demo2010, by = join_by(DE_ID))
  
```

Import all available records for calculating HOMA1 estimates between 2020-2024.

```{r}

import_dir <- here::here("exports/UAB/raw")
glucose2020 <- read_csv(file.path(import_dir, "glucose.csv"), show_col_types = F)
cpeptide2020 <- read_csv(file.path(import_dir, "cpeptide.csv"), show_col_types = F)
insulin2020 <- read_csv(file.path(import_dir, "insulin.csv"), show_col_types = F)
birthday2020 <- read_csv(file.path(import_dir, "birthdays.csv"), show_col_types = F)
demo2020 <- read_csv(file.path(import_dir, "patient_demo.csv"), show_col_types = F)
```

Filter data to only records containing glucose, insulin and c-peptide values measured at the same time.

```{r}
data2020 <- glucose2020 %>%
  inner_join(cpeptide2020, by = c("DE_ID" = "DE_ID",
                              "glucose_visit_number" = "cpeptide_visit_number", 
                              "glucose_date_time" = "cpeptide_date_time")) %>%
  inner_join(insulin2020, by = c("DE_ID" = "DE_ID",
                              "glucose_visit_number" = "insulin_visit_number", 
                              "glucose_date_time" = "insulin_date_time")) %>%
  inner_join(birthday2020, by = join_by(DE_ID)) %>%
  inner_join(demo2010, by = join_by(DE_ID))
```

Combine data and calculate HOMA1 estimates from both insulin and c-peptide. Records with glucose values less than or equal to 3.5 mM were excluded since they result in negative or infinity HOMA1-B values.

```{r}

data_uab <- data2010 %>%
  bind_rows(data2020)

data_uab <- data_uab %>% 
  # filter to HOMA range
  filter(glucose > 3.5 & glucose <= 25 &
           cpeptide >= 0.2 & cpeptide <= 3.5 & 
           insulin >= 20 & insulin <= 400) %>%
  mutate(cpeptide_homa1b = (20 * 6 * cpeptide) / (glucose - 3.5),
         cpeptide_homa1ir = (glucose * 6 * cpeptide) / 22.5,
         insulin_homa1b = (20 * insulin/6) / (glucose - 3.5),
         insulin_homa1ir = (glucose * insulin/6) / 22.5,
         source = "uab", # label data source
         age = as.numeric(date(glucose_date_time) - birth_date)/365) %>%
  filter(age >= 18) # keep only adults

```

Unify demographic features with NHANES.

```{r}
data_uab <- data_uab %>%
  mutate(race = relevel(factor(case_when(race == "White" ~ "Non-Hispanic White",
                                         race == "Black or African American" ~ "Non-Hispanic Black",
                                         race == "Hispanic or Latino" ~ "Hispanic or Latino",
                                         .default = "Other Race - Including Multi-Racial")), ref = "Non-Hispanic White"),
         gender = case_when(gender == "F" ~ "Female",
                            gender == "M" ~ "Male"))
```

## NHANES

Import insulin, c-peptide, and glucose data. Due to instrument changes between 1999-2002 and 2003-2004 from the Pharmacia to the Tosoh, insulin values from 2003-2004 were regressed backwards to match the 1999-2002 values.

```{r}
nhanes1999 <- nhanes("LAB10AM") %>% as_tibble()
nhanes2001 <- nhanes("L10AM_B") %>% as_tibble()
nhanes2001_2 <- nhanes("L10_2_B") %>% as_tibble() %>%
  rename(LBXGLUSI = LB2GLUSI,
         LBXCPSI = LB2CPSI,
         LBXINSI = LB2INSI)
nhanes2003 <- nhanes("L10AM_C") %>% as_tibble() %>%
  rename(LBXGLUSI = LBDGLUSI) %>%
  mutate(LBXINSI =  (0.96006 * LBXIN + 3.23663) * 6) # regress backwards and convert to SI units

nhanes_data <- bind_rows(list(nhanes1999, nhanes2001, nhanes2001_2, nhanes2003)) %>%
  # filter to HOMA range
  filter(LBXGLUSI > 3.5 & LBXGLUSI <= 25 &
           LBXCPSI >= 0.2 & LBXCPSI <= 3.5 & 
           LBXINSI >= 20 & LBXINSI <= 400) %>% 
  mutate(glucose = LBXGLUSI,
         cpeptide_homa1b = (20 * 6 * LBXCPSI) / (LBXGLUSI - 3.5),
         cpeptide_homa1ir = (LBXGLUSI * 6 * LBXCPSI) / 22.5,
         insulin_homa1b = (20 * LBXINSI/6) / (LBXGLUSI - 3.5),
         insulin_homa1ir = (LBXGLUSI * LBXINSI/6) / 22.5,
         source = "nhanes") %>% # label data source
  select(SEQN, glucose, cpeptide_homa1b, cpeptide_homa1ir, insulin_homa1b, insulin_homa1ir, source) %>%
  drop_na()
```

Import demographic data for participants' ages and insulin use status.

```{r}
table_names <- c("DEMO", "DIQ") # for survey age , sex, race, and insulin status

table_suffixes <- c("nosuffix", "B", "C") # for 1999-2000, 2001-2002, 2003-2004

nhanes_tables <- list()

for (s in table_suffixes){
  nhanes_tables[[s]] <- list()
  
  table_codes <- purrr::map_vec(table_names, paste, s, sep = "_")
  table_codes <- map(table_codes, ~str_remove(string = .x, pattern = "_nosuffix$"))
  
  nhanes_tables[[s]] <- map(table_codes, nhanes) %>% 
    map(., as_tibble)
  
  names(nhanes_tables[[s]]) <- table_names
}
```

Merge demographic data across cycles.

```{r}
cycle_tables <- list()

for (s in table_suffixes){
  cycle_tables[[s]] <- nhanes_tables[[s]]$DEMO %>% mutate(cycle = s)
  
  for (t in table_names[table_names != "DEMO"]){
    
    cycle_tables[[s]] <- cycle_tables[[s]] %>% 
      full_join(nhanes_tables[[s]][[t]]) #%>% select(-one_of("WTPH2YR")) # remove phebotomy weight, causes duplication issues
  }
}

merged_nhanes <- bind_rows(cycle_tables)
```

Join demographic data with HOMA data.

```{r}
nhanes_data <- nhanes_data %>% inner_join(merged_nhanes, by = join_by(SEQN)) %>%
  rename(DE_ID = SEQN) # rename ID to match UAB data
```

Unify demographic features with NHANES.

```{r}
nhanes_data <- nhanes_data %>%
  mutate(race = relevel(factor(case_when(RIDRETH1 == "Mexican American" ~ "Hispanic or Latino",
                          RIDRETH1 == "Other Hispanic" ~ "Hispanic or Latino",
                          .default = RIDRETH1)), ref = "Non-Hispanic White")) %>%
  rename(gender = RIAGENDR)
```

Remove NHANES participants under 18 and remove participants currently using insulin.

```{r}
nhanes_data <- nhanes_data %>% filter(RIDAGEYR >= 18) %>% filter(DIQ050 != "Yes")
```

## Combine data

```{r}
data <- bind_rows(list(data_uab, nhanes_data))
```

Combined data contains 3003 HOMA estimates pairs from 2720 UAB patients and 6510 HOMA estimate pairs from 6331 NHANES participants.

Set some data aside for testing.

```{r}

test_data <- data %>% slice_sample(prop = 0.2, replace = F)

data <- data %>% anti_join(test_data, by = join_by(DE_ID))

```

Training data contains 2229 HOMA estimate pairs from 2112 UAB patients and 5175 HOMA estimate pairs from 5060 NHANES partipants. 
Test data contains 626 HOMA estimate pairs from 608 UAB patients and 1276 HOMA estimate pairs from 1271 NHANES participants.

# Linear regression

HOMA1 estimate values appear approximately log-normal and linearly correlated after log-transformation. UAB data seems to contain lower HOMA1-B and higher HOMA1-IR values, likely since UAB data contains only patients with diabetes while most NHANES participants do not have diabetes.

```{r}

grid.arrange(# HOMA-B
             ggplot(data, mapping = aes(x = cpeptide_homa1b)) + geom_histogram(),
             ggplot(data, mapping = aes(x = log(cpeptide_homa1b))) + geom_histogram(),
             ggplot(data, mapping = aes(x = insulin_homa1b)) + geom_histogram(),
             ggplot(data, mapping = aes(x = log(insulin_homa1b))) + geom_histogram(),
             # HOMA-IR
             ggplot(data, mapping = aes(x = cpeptide_homa1ir)) + geom_histogram(),
             ggplot(data, mapping = aes(x = log(cpeptide_homa1ir))) + geom_histogram(),
             ggplot(data, mapping = aes(x = insulin_homa1ir)) + geom_histogram(),
             ggplot(data, mapping = aes(x = log(insulin_homa1ir))) + geom_histogram(),
             # correlation
             ggplot(data, mapping = aes(x = log(insulin_homa1b), y = log(cpeptide_homa1b), color = source)) +
               geom_smooth(method = "lm") + geom_density_2d(),
             ggplot(data, mapping = aes(x = log(insulin_homa1ir), y = log(cpeptide_homa1ir), color = source)) + 
               geom_smooth(method = "lm") + geom_density_2d(),
             ncol = 2)
```

Confirmation linear correlation.

```{r}
cor.test(log(data$insulin_homa1b), log(data$cpeptide_homa1b))
cor.test(log(data$insulin_homa1ir), log(data$cpeptide_homa1ir))
```


## HOMA1-B regression

Build linear regression model using log-transformed HOMA1 values. Since some patients have more than 1 value, DE_ID was included as a random effect to account for the nested structure of the data. 

```{r}
homa1b_model <- lmer(log(cpeptide_homa1b) ~ log(insulin_homa1b) + race + gender + (1|DE_ID), data)

homa1b_summary <- summary(homa1b_model)

print(homa1b_summary)

print(r.squaredGLMM(homa1b_model))

print(homa1b_summary$coefficients)
```

Inspect regression.

```{r}
plot(homa1b_model)
```

## HOMA1-IR regression

Build linear regression model using log-transformed HOMA1 values. DE_ID were included as factors so that regression won't be weighed towards patients with multiple HOMA1 estimate values. 

```{r}
homa1ir_model <- lmer(log(cpeptide_homa1ir) ~ log(insulin_homa1ir) + race + gender + (1|DE_ID), data)

homa1ir_summary <- summary(homa1ir_model)

print(homa1ir_summary)

print(r.squaredGLMM(homa1ir_model))

print(homa1ir_summary$coefficients)
```

Inspect regression.

```{r}
plot(homa1ir_model)
```

# Inspect model

Calculate predicted estimate values from linear regression. Random effects were ignored since we only want a general conversion for a "population average".

```{r}
data <- data %>%
  mutate(homa1b_pred = exp(predict(homa1b_model, newdata = data, re.form = NA)), # re.form = NA to specify no random effects
         homa1ir_pred = exp(predict(homa1ir_model, newdata = data, re.form = NA)))
```

Compare mean absolute error.

```{r}

# calculate absolute difference
data <- data %>% 
  mutate(homa1b_ad = abs(cpeptide_homa1b - insulin_homa1b),
         homa1ir_ad = abs(cpeptide_homa1ir - insulin_homa1ir),
         homa1b_fit_ad = abs(cpeptide_homa1b - homa1b_pred),
         homa1ir_fit_ad = abs(cpeptide_homa1ir - homa1ir_pred))

abdif_list <- list("homa1b_ab_dif" = summary(data$homa1b_ad) %>% as.list() %>% as_tibble(),
                   "homa1ir_ab_dif" = summary(data$homa1ir_ad) %>% as.list() %>% as_tibble(),
                   "homa1b_fitted_ab_dif" = summary(data$homa1b_fit_ad) %>% as.list() %>% as_tibble(),
                   "homa1ir_fitted_ab_dif" = summary(data$homa1ir_fit_ad) %>% as.list() %>% as_tibble())


abdif_summary <- bind_rows(abdif_list) %>%
  mutate(model = names(abdif_list),
         pairedTTest = c(NA, NA, 
                         t.test(data$homa1b_fit_ad, data$homa1b_ad, paired = T)$p.value,
                         t.test(data$homa1ir_fit_ad, data$homa1ir_ad, paired = T)$p.value)) %>%
  relocate(model)

```

Regression does appear to shift insulin HOMA1 estimate values closer to c-peptide HOMA1 estimate values.

```{r}
# graphing parameters
base_size <- 7
size = 0.1

b_cpeptide_insulin <- ggplot(data, mapping = aes(x = log(cpeptide_homa1b), y = log(insulin_homa1b))) + 
  theme_classic(base_size = base_size) +
  geom_point(size = size) + 
  geom_density_2d() + 
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(x = expression(ln("HOMA1-B"["c-peptide"])), 
       y = expression(ln("HOMA1-B"["insulin"])),
       color = "Data source") + 
  theme(plot.margin = margin(t = 6))

b_cpeptide_fitted <- ggplot(data, mapping = aes(x = log(cpeptide_homa1b), y = log(homa1b_pred))) + 
  theme_classic(base_size = base_size) +
  geom_point(size = size) + 
  geom_density_2d() + 
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(x = expression(ln("HOMA1-B"["c-peptide"])), 
       y = expression(ln("Scaled HOMA1-B"["insulin"]))) + 
  theme(plot.margin = margin(t = 6))

ir_cpeptide_insulin <- ggplot(data, mapping = aes(x = log(cpeptide_homa1ir), y = log(insulin_homa1ir))) + 
  theme_classic(base_size = base_size) +
  geom_point(size = size) + 
  geom_density_2d() +
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(x = expression(ln("HOMA1-IR"["c-peptide"])), 
       y = expression(ln("HOMA1-IR"["insulin"]))) + 
  theme(plot.margin = margin(t = 6)) 

ir_cpeptide_fitted <- ggplot(data, mapping = aes(x = log(cpeptide_homa1ir), y = log(homa1ir_pred))) + 
  theme_classic(base_size = base_size) +
  geom_point(size = size) + 
  geom_density_2d() +
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(x = expression(ln("HOMA1-IR"["c-peptide"])), 
       y = expression(ln("Scaled HOMA1-IR"["insulin"]))) + 
  theme(plot.margin = margin(t = 6))  

grid.arrange(b_cpeptide_insulin,
             b_cpeptide_fitted,
             ir_cpeptide_insulin,
             ir_cpeptide_fitted,
             ncol = 2)

```

Save plots.

```{r}
export_dir <- here::here("exports", "HOMA_regression")

if (!dir.exists(export_dir)) {
  dir.create(export_dir, recursive = T)
}

# graphing parameters
dpi <- 600
width <- 1.5
height <- 1.5
units <- "in"

# save plots
ggsave(file.path(export_dir, "homa_regression_b_cpep_ins.tiff"), b_cpeptide_insulin,
       width = width, height = height, units = units, dpi = dpi)
ggsave(file.path(export_dir, "homa_regression_b_cpep_fit.tiff"), b_cpeptide_fitted,
       width = width, height = height, units = units, dpi = dpi)
ggsave(file.path(export_dir, "homa_regression_ir_cpep_ins.tiff"), ir_cpeptide_insulin,
       width = width, height = height, units = units, dpi = dpi)
ggsave(file.path(export_dir, "homa_regression_ir_cpep_fit.tiff"), ir_cpeptide_fitted,
       width = width, height = height, units = units, dpi = dpi)

# save absolute difference
write_csv(abdif_summary, file.path(export_dir, "homa_regression_ab_dif.csv"))
```

# Test model on hold-out data

```{r}
test_data <- test_data %>%
  mutate(homa1b_pred = exp(predict(homa1b_model, newdata = test_data, re.form = NA)),
         homa1ir_pred = exp(predict(homa1ir_model, newdata = test_data, re.form = NA))) # re.form = NA to specify no random effects
```

```{r}
# graphing parameters
base_size <- 7
size = 0.1

b_cpeptide_insulin <- ggplot(test_data, mapping = aes(x = log(cpeptide_homa1b), y = log(insulin_homa1b))) + 
  theme_classic(base_size = base_size) +
  geom_point(size = size) + 
  geom_density_2d() + 
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(x = expression(ln("HOMA1-B"["c-peptide"])), 
       y = expression(ln("HOMA1-B"["insulin"]))) + 
  theme(plot.margin = margin(t = 6))

b_cpeptide_fitted <- ggplot(test_data, mapping = aes(x = log(cpeptide_homa1b), y = log(homa1b_pred))) + 
  theme_classic(base_size = base_size) +
  geom_point(size = size) + 
  geom_density_2d() + 
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(x = expression(ln("HOMA1-B"["c-peptide"])), 
       y = expression(ln("Scaled HOMA1-B"["insulin"]))) + 
  theme(plot.margin = margin(t = 6))

ir_cpeptide_insulin <- ggplot(test_data, mapping = aes(x = log(cpeptide_homa1ir), y = log(insulin_homa1ir))) + 
  theme_classic(base_size = base_size) +
  geom_point(size = size) + 
  geom_density_2d() +
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(x = expression(ln("HOMA1-IR"["c-peptide"])), 
       y = expression(ln("HOMA1-IR"["insulin"]))) + 
  theme(plot.margin = margin(t = 6)) 

ir_cpeptide_fitted <- ggplot(test_data, mapping = aes(x = log(cpeptide_homa1ir), y = log(homa1ir_pred))) + 
  theme_classic(base_size = base_size) +
  geom_point(size = size) + 
  geom_density_2d() +
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(x = expression(ln("HOMA1-IR"["c-peptide"])), 
       y = expression(ln("Scaled HOMA1-IR"["insulin"]))) + 
  theme(plot.margin = margin(t = 6))  

grid.arrange(b_cpeptide_insulin,
             b_cpeptide_fitted,
             ir_cpeptide_insulin,
             ir_cpeptide_fitted,
             ncol = 2)

```

```{r}
# calculate absolute difference
test_data <- test_data %>% 
  mutate(homa1b_ad = abs(cpeptide_homa1b - insulin_homa1b),
         homa1ir_ad = abs(cpeptide_homa1ir - insulin_homa1ir),
         homa1b_fit_ad = abs(cpeptide_homa1b - homa1b_pred),
         homa1ir_fit_ad = abs(cpeptide_homa1ir - homa1ir_pred))

abdif_list <- list("homa1b_ab_dif" = summary(test_data$homa1b_ad) %>% as.list() %>% as_tibble(),
                   "homa1ir_ab_dif" = summary(test_data$homa1ir_ad) %>% as.list() %>% as_tibble(),
                   "homa1b_fitted_ab_dif" = summary(test_data$homa1b_fit_ad) %>% as.list() %>% as_tibble(),
                   "homa1ir_fitted_ab_dif" = summary(test_data$homa1ir_fit_ad) %>% as.list() %>% as_tibble())

abdif_summary <- bind_rows(abdif_list) %>%
  mutate(model = names(abdif_list),
         pairedTTest = c(NA, NA, 
                         t.test(test_data$homa1b_fit_ad, test_data$homa1b_ad, paired = T)$p.value,
                         t.test(test_data$homa1ir_fit_ad, test_data$homa1ir_ad, paired = T)$p.value)) %>%
  relocate(model)
```

Save graphs and summary differences.

```{r}

# save plots
ggsave(file.path(export_dir, "homa_regression_b_cpep_ins_test.tiff"), b_cpeptide_insulin,
       width = width, height = height, units = units, dpi = dpi)
ggsave(file.path(export_dir, "homa_regression_b_cpep_fit_test.tiff"), b_cpeptide_fitted,
       width = width, height = height, units = units, dpi = dpi)
ggsave(file.path(export_dir, "homa_regression_ir_cpep_ins_test.tiff"), ir_cpeptide_insulin,
       width = width, height = height, units = units, dpi = dpi)
ggsave(file.path(export_dir, "homa_regression_ir_cpep_fit_test.tiff"), ir_cpeptide_fitted,
       width = width, height = height, units = units, dpi = dpi)

# save absolute difference
write_csv(abdif_summary, file.path(export_dir, "homa_regression_ab_dif_test.csv"))
```

Save testing data for downstream testing.

```{r}
write_csv(test_data, file.path(export_dir, "homa_regression_test_data.csv"))
```

Save models.

```{r}
saveRDS(homa1b_model, file.path(export_dir, "homa1b_model.RDS"))
saveRDS(homa1ir_model, file.path(export_dir, "homa1ir_model.RDS"))
```

# Session information

```{r}
sessionInfo()
```

